---
title: "Econ 343 Ch 6 HW C8 C 13 Lastname, Firstname"
output:
  pdf_document: 
    toc: yes
  html_document:
    df_print: paged
  word_document: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(scipen=999)
options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```

Instructions: 

- Save your commands and show all work for the Cengage computer exercises
- the variable names should make it obvious what you are computing

- Each problem should begin by providing the context - that is, the steps of empirical analysis:

1. Economic Question
2. Economic Model
3. Econometric Model
4. Data
5. Analysis

- Then begin inserting the parts of the Cengage problem
- perform all calcuations in R chunks
- present all results in the write-up below the r chunks - using r inline to recall any estimated or calculated results

- For each part in the problem:
  - Include the following sections as needed:
  - _Description_ Explain what the purpose of that part of the question is
  - _Estimated Equation_
    - Unless indicated otherwise, write out the estimated regression equation using R inline.
    - Use the `\underset{}{}` command to put the standard errors under each of the slope coefficients in the equation

  - _Results_ 
    - Interpret what the estimated coefficient means, put into real terms when possible and/or helpful
    - Present the results and conclusion of any hypothesis tests


# Ch. 6 C8 House Price II

## Part (i) Regression

``` {r}
load(???)
mydata <- Data
lm6.8i <- lm(???)

```

### Prediction

```{r}
# Review Ch 6 Lecture for how to use the predict() command

# An example for how to do the first problem is below. You will
# need to change the values as needed to match your problem on Cengage

xvals <- data.frame(lotsize = 10000,sqrft = 2000, bdrms = 3)

ypred1 <- predict(lm6.8i,xvals)

```




## Part (ii) CI for predicted $\hat{y}$:

Cengage walks you through how to "trick"
the regression to give you the confidence interval
around the prediction in part (i). This is similar to using algebra to "trick" the regression to provide the results of 
a linear hypothesis test in Ch 4. 

Although I have you practice the "trick" approach to answer the Cengage problem, I also give you the one-line of code that you can use to check your answer and use instead. 

The five steps to compute the CI for a prediction by hand: 

1. Run a new regression
2. Obtain the $\hat\beta$ for the constant 
3. Obtain the $se_{\hat\beta}$ from the constant
4. Look up the *tvalue* for the d.f. of the model
5. Plug these numbers into the CI formula and calculate the lower and upper CI by hand. 

Use following code chunk to give you the answers you need for part (ii).

### Method 1: Calculate CI by hand 

1. Trick the regression to provide the *se* of the point-estimate.

```{r}

# Again, I have given you an example for how to do this, but you need
# to change the values as needed
# Subtract the input values for each variable from
# the variable in the regression.
lm6.8ii <- lm(price ~ 
                I(lotsize - 10000) +
                I(sqrft - 2000) +
                I(bdrms - 3), data = mydata)

stargazer(lm6.8ii, type = "text")

# The point-estimate and se of the prediction is from the Constant!


# Obtain the critical value
alpha <- 0.05
cval <- abs(qt(alpha/2,???))

# Calculate the CI


```

2. Obtain the point-estimate for the price of a house with these
characteristics from the constant in
the estimated regression $\hat\theta$ = `r point.estimate`,

3. Obtain the standard error of the estimate from the saved results of the regression $se_{\hat\theta}$ = `r se`.

4. Find the *critical value* for a 95% confidence interval with df = 84 (n-k-1) (using  `qt(alpha/2,df)` command in `R` *C.V.* = `r cval`.


5. Use the standard confidence interval formula to calculate the CI:
$$\hat\theta \pm
t_{\alpha/2,df=n-k-1} * se_{\hat\beta_{Intercept}}$$

$$`r point.estimate` \pm `r cval` * `r se`$$
Write out the confidence interval for expected price (in thousands) for houses with these characteristics as **[$`r lowerci`, $`r upperci`]**.

\newpage

In *R*, you can execute all of the above steps with just one command:

### Method 2: Use predict() to obtain confidence interval

In *R*, you can simply use `predict()` to accomplish all of the above steps in one command.

```{r}
# Review Ch 6 R Lecture for how to use interval = "confidence" in the predict command



# Review Lecture 6 for how to save the predict() results to use in r inline.


```



## Part (iv) CI for single prediction (WIDE prediction)

In part (iii) the confidence interval was for the predicted selling price for a house with certain characteristics.
Part (iv) asks for the CI for the price of a particular house. 

Simply use `interval = "predict"` in the predict() to get the wider CI.

```{r}


```



\newpage

# Ch. 6 C13 School Expenditures

## Part (i) Regression

``` {r}
load(???)
mydata <- Data
lm6.13i <- ???

```


## Part (ii) Compare ranges of $y$ and $\hat{y}$

```{r}
## Use mutate with the fitted () command to add the fitted values 
# as a new variable in the dataset. 

mydata <- mydata %>% 
  mutate(yhat = fitted(lm6.13i))   


```



## Part (iii) Find largest residual

```{r}
# Residuals

# Use the mutate command with the resid() command 
# to add the residuals as a new variable to the dataset
mydata <- mydata %>%
  mutate(resid1 = resid(lm6.13i))


```


## Part (iv) Joint Significance test

```{r}
# Run the new regression
lm6.13iv <- ???

# Use linearHypothesis() to conduct joint significance tests


# Review Ch 4 R Lecture for saving test results as variables to use 
# in r inline in the write-up below


```



## Part (v) Scale()

```{r}
# CAUTION: When running a scaled regression, don't
# forget to add the "0" to indicate no constant.
lm6.13v <- lm(???)

stargazer(lm6.13v,type = "text")
```



