---
title: "Econ 343 Ch 3 C2 Report Lastname, firstname"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```
# Ch 03 Computer Assignment C2 - Housing Prices Report


Instructions: In addition to using the results for the homework on Cengage, 
you will write up the analysis as a small "report." Write of the results for each part of the homework in a sentence or two here. Show all work by doing
math in R chunks and using inline R to refer to the result in the write-up. 

To get you started, Part (i) is
done for you in the format of a miniature econometric report. However,
this same
format is followed for even the most sophisticated econometric analysis.

```{r}
# Download the dataset from Cengage and then upload the dataset to Rstudio
# change the location of the dataset in the load command, as needed

load("~/ECON 343/Homework/Data/hprice1_v1.RData")
mydata <- Data

stargazer(mydata, type = "text")

# If the above stargazer() command Use the as.data.frame() command to ensure the data can be used by stargazer()
stargazer(as.data.frame(mydata), type = "text")


```

# Part (i)
```{r}
# Enter the R commands you need to answer this part of the question
# Do this as needed for all of the subsequent parts

lm1 <- lm(price ~ bdrms + sqrft ,data = mydata)
bhat1 <- coef(lm1)

stargazer(lm1,type = "text",digits = 2)

```

# Basic Format of Empirical Analysis

**1. Economic Question**

How much does the size and number of bedrooms of a house affect its price?

**2. Economic Model**

$$price = f(size, location, location, location)$$

**3. Econometric Model**

$$price = \beta_0 + \beta_1 sqrft + \beta_2 bdrms + u $$

**4. Data**

To examine this question, house price, sq. ft, number of bedrooms, and other 
information was gathered on 83 homes from the real estate pages of the *Boston Globe* newspaper in 1990. 

---

**5. Results**

House price(in thousands) was regressed on sq. footage and number of bedrooms. The
estimated equation is:

$$\widehat{price} = 
  `r round(bhat1["(Intercept)"],2)`
+ `r round(bhat1["sqrft"],2)`sqrft 
+ `r round(bhat1["bdrms"],2)`bdrms $$

## Part (ii)

\$ this is a test `$`

## Part (iii)


## Part (iv)


## Part (v)


## Part (vi)
