---
title: "Econ 343 Ch 2 HW Lastname, firstname"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```
# Ch 02 Computer Assignment #C2.1

### New commands used in this problem:
- lm()
- stargazer()
- coef()
- fitted()
- computing R2 by hand

### Rmarkdown syntax used in this write-up sections below:

- Use a backslash before a dollar sign or percent sign

```{r}
#Load the dataset manually because each Cengage assignment is a different version
load("~/20SU ECON 343/Homework/401k_v4.RData")
mydata <- Data  
# as_tibble(mydata)  # Uncomment this command the first time the data is loaded 
                     # to check that the data looks as expected.
```

## Part (i)
```{r}
avg_prate <- mean(mydata$prate)
avg_mrate <- mean(mydata$mrate)
```
The average participation rate in 401k plans in this dataset is **`r avg_prate`**. The average match rate is **`r avg_mrate`**.

## Part (ii)
```{r}
lm.1 <- lm(prate ~ mrate,data=mydata) 
stargazer(lm.1,type="text")

bhat <- coef(lm.1)   # create "bhat" as a vector of the estimated coefficients
                     # to make it easier to refer to individual 
                     # estimated coefficients inline
```


The coefficient on match rate (*mrate*) is **`r bhat["mrate"]`**. 

**Interpretation**: for every
\$1.00 that a company matches of an employee's \$1.00 contribution to a 401k, the participation
rate in 401k plans among employees increases on average by **`r bhat["mrate"]`** percentage points.

## Part (iii)
For a \$0.30 increase in how much the company contributes for every $1.00 the
employee contributes, the participation
rate of employees increases on average by **`r .3 * bhat["mrate"]`** percentage points.

## Part (iv)
A company that has a matching rate of 1.5 is expected to have **`r bhat["(Intercept)"] + (1.5 * bhat["mrate"])`** participation rate in
401k plans by its employees.

## Part (v)
The *R-squared* value for this estimated regression model is 
**`r var(fitted(lm.1)) / var(mydata$prate)`**.
