---
title: "Econ 343 Ch 2 HW Lastname, firstname"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```
# Ch 02 Computer Assignment #C2.1

### R commands used in the r chunks below:
- lm()
- stargazer()
- coef()
- fitted()
- computing R2 by hand

### Rmarkdown syntax used in this write-up sections below:

- Use a backslash before a dollar sign or percent sign

CAUTION! This Rmarkdown file will not knit until you complete the commands
as indicated below with **TODO:**.

```{r}
## Upload your version of the data from Cengage to Rstudioand point the load command
## to its location and give it the correct name

load("~/ECON 343-yourname/Homework/401k_v?.RData")
mydata <- Data  

as_tibble(mydata)  
```

## Part (i)
```{r}
# TODO: replace the# "???" below with the correct command and uncomment the line
#avg_prate <- ???
#avg_mrate <- ???
```
The average participation rate in 401k plans in this dataset is **`r avg_prate`**. The average match rate is **`r avg_mrate`**.

## Part (ii)
```{r}
# TODO: uncomment the correct lm() model below
#lm.1 <- lm(prate ~ mrate,data=mydata)   # regresses prate on mrate
#lm.1 <- lm(mrate ~ prate,data=mydata)   # regresses mrate on prate

stargazer(lm.1,type="text")   # presents nicely formatted regression results

bhat <- coef(lm.1)   # create "bhat" as a vector of the estimated coefficients
                     # to make it easier to refer to individual 
                     # estimated coefficients inline
```


The coefficient on matching rate (*mrate*) is **`r bhat["mrate"]`**. 

**Interpretation**: for every
\$1.00 that a company matches of an employee's \$1.00 contribution to a 401k,
the participation
rate of employees
increases on average by **`r bhat["mrate"]`** percentage points.

## Part (iii)
```{r}

result_1 <- 0.30 * bhat["???"]

```

For every \$0.30 increase in how much the company contributes for every \$1.00 the
employee contributes, the participation
rate of employees increases on average by **`r result_1`**.

**Explanation**:
- The interpretation formula for estimated coefficients is: 
**For every 1 unit increase in X, Y increases on average by **$\hat{\beta_?}$.**


- So, for a 0.30 unit increase in X, the average increase in Y will be 0.30 multiplied by
the value of the estimated coefficient on X. 

- 0.30 times $\hat{\beta}_{mrate}$ = 0.30 times `r bhat["mrate"]` = **`r result_1`** percentage points. 

## Part (iv)

```{r}

result_2 <- bhat["(Intercept)"] + (1.5 * ????["mrate"])

```


A company that has a matching rate of 1.5 is expected to have **`r result_2`** participation rate in
401k plans by its employees, based on the estimated regression equation below:

$$\widehat{prate} = \hat{\beta_0} + \hat{\beta_1} * \textbf{mrate}$$
$$\widehat{prate} = `r bhat["(Intercept)"]` + `r bhat["mrate"]` * \textbf{mrate}$$

$$\widehat{prate} = `r bhat["(Intercept)"]` + `r bhat["mrate"]` * 1.5$$

$$\widehat{prate} = `r bhat["(Intercept)"] + bhat["mrate"] * 1.5`$$

## Part (v)
```{r}
actual_y     <- mydata$prate
fitted_y     <- fitted(lm.1)
var_fitted_y <- var(fitted_y)
var_actual_y <- ???
r_sq_by_hand <- ???
```

The *R-squared* value (calculated by hand) for this estimated regression model is 
**`r ______`**. 

Explanation:
The *R-squared* value for an estimated regression model is calculated as the percent
of the total variance in the dependent variable that is explained by the regression model.
One way to calculate this, is to divide the variance of the fitted values by the variance of the actual values.

The total variance in the actual values of the dependent variable is
$$var(y) = var(prate) = `r _____`$$ 

The variance of the fitted values of the dependent variable is
$$var(\hat{y}) = var(fitted) = `r _______`$$

The variance 
explained by the regression model is thus: 
$$\frac{var(fitted)}{var(actual)} = \frac{`r _____`}{`r _____`} = `r _______`$$.

