---
title: "Econ 343 Ch 7 HW C8 Lastname, firstname  "
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(sjmisc)
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data
library(kableExtra)

options(scipen=999)
options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```

**Instructions:**

- The primary learning objective for this HW is interpreting dummy variables.
- Save your commands and show all work for the Cengage computer exercises
- the variable names should make it obvious what you are computing

- The problem should begin by providing the context - that is, the steps of empirical analysis:

1. Economic Question
2. Economic Model
3. Econometric Model
4. Data
5. Analysis

- Then begin inserting the parts of the Cengage problem
- perform all calcuations in R chunks
- present all results in the write-up below the r chunks - using r inline to recall any estimated or calculated results

- For each part in the problem:
  - Include the following sections as needed:


_Description_ Explain what the purpose of that part of the question is

_Estimated Equation_ FOR THIS ASSIGNMENT - YOU DO NOT NEED TO WRITE OUT the estimated regression equation using R inline. BUT, always display the regression results using stargazer(). Display regression results side-by-side as indicated below.

_Results_ 
    - Interpret what the estimated coefficient on `white` means after every regression
    - Present the results and conclusion of any hypothesis tests

# Ch. 7 C8 Mortgage Approval Rates Linear Probability Model (LPM)

## Part (i)

Nothing required here.


## Part (ii)

``` {r}
# Run the appropriate regression.

lm7C8.ii <- 
  

```



## Part (iii)

``` {r}
# Run the appropriate regression.

lm7C8.iii <-

```

## Part (iv)

``` {r}
# Run the appropriate regression.

lm7C8.iv <-

# Calculate the confidence interval as specified in Cengage.

```





## Part (v)

``` {r}
# Run the appropriate regression.
lm7C8.v <-
  

  
  
# HINT: Check out the confint() command in the stats package for this problem.
  

```



# I. LPM: Check predicted probabilities

In this section of your HW, you are asked to go beyond the questions in the Cengage HW. For this section, follow the steps in the Ch 7 lecture to examine the validity of the predicted probabilities of the last model that was estimated (lm7.8v). That is, are any of the predicted probabilities negative or greater than 1? 

Discuss the results using the same format that the results are discussed in the **Write-up** portion of the explantion in the lecture.

**Additional Instructions:**

- Use the full dataset `loanapp`, not the Cengage version.

- Eliminate any observations (rows) that have missing data (`NA`) before running the regression

- These two steps are accomplished for you in the first command in the r chunk below that creates the dataset: mydata.nomissings

- Be sure to use `mydata.nomissings` for the rest of the HW assignment

```{r}

mydata.nomissings <- na.omit(loanapp)

lm1.I <- lm( ???, data=mydata.nomissings)


```





# II. LPM Compare actual to predicted probabilities

In this section of your HW, you are asked to go beyond the questions in the Cengage HW. For this section of your HW, follow the steps in the Ch 7 lecture to examine how well the last model that was estimated (lm7.8v) does at predicting mortgage approvals.

Discuss the results using the same format that the results are discussed in the **Write-up** portion of the explantion in the lecture.

```{r}
# To fix knitting error, use "the "split_table_to_df() %>%   kable(booktabs = T)"
# after every cro_() (or related) command

# Example:

# cro_cpct(mydata.nomissings$approve, mydata.nomissings$yhat_dummy) %>%
#  split_table_to_df() %>%
#  kable(booktabs = T)

```

