---
title: "Econ 343 Ch 10 HW **student name**"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```

**Instructions:**

- The primary learning objective for this HW is interpreting time series regressions.
- Answer all questions that are listed after an R chunk.
- Each question is worth 10 points (Total 50 pts)

- The R commands you need to answer the Cengage HW are provided. 
- Don't forget to load your own **unique dataset from Cengage**.

# R commands used in this HW

- ts()
- dynlm()
- L()


# Ch. 10 C7 - Consumption and Income

## Read in the CENGAGE data

``` {r}

# load("consump_v1.RData")      # <<<<< change this to your own Cengage version 
# mytsdata <- ts(Data,start = 1959) 

mytsdata <- ts(consump,start = 1959) 


                                     
as_tibble(mytsdata)

```

## Part (i)
``` {r}

lm.i <- dynlm()

summary(lm.i)

stargazer(lm.i,type = "text")

```

## Part (ii) add lag
``` {r}

# For full credit, use the "L()" command to create the lagged variable in the regression

lm.ii <- dynlm()

summary(lm.ii)

stargazer(lm.i,lm.ii,type = "text")


```


## Part (iii) additional regressor
``` {r}

lm.iii <- dynlm()

summary(lm.iii)

stargazer(lm.i,lm.ii,lm.iii,type = "text")


```
