---
title: "Econ 343 Ch 1 HW Lastname, Firstname"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```
# Ch 01 Computer Assignment #C1.3
## Using R for your first computer homework assignment.
### Commands used in this problem:
- max()
- min()
- filter()
- as_tibble()
- as.numeric()
- count()
- mean()
- sd()
- cor()
- using R as a calculator

INSTRUCTIONS: Fill in the blanks or "???" in the r-chunks and corresponding write-ups below. Some of the questions have already been answered for you to help you get started.

# Load the data

- First upload your own version of meap01 from Cengage, for example "meap01_v5.RData" 
- Change the filename in the load statement below as needed

```{r}
#Load the meap01_v?? data manually since it is not in the wooldridge package
load("~/Econ-343-newname/Homework/meap01_v5.RData")
mydata <- Data
as_tibble(mydata)
```


## Part (i) Simple Descriptions

Save the results of the min() and max() commands to refer to them in the write-up
below using inline r. 

```{r}

min_math4 <- min(mydata$math4)
max_math4 <- max(mydata$math4)

```
[NOTE: It is common practice to italicize names of variables, and
**bold** numerical results, as in the following sentence.]

The lowest *math4* passing rate is **`r min_math4`**. 

The highest value of *_____* is **______________**.

# Part (ii)

## How many have perfect math4 scores?

```{r}

mydata.perfectmath <- filter(mydata,math4==100)
n_perfectmath <- as.numeric(count(mydata.perfectmath))
```

The number of schools with a perfect pass rate on the math test is
**`r ______ `**.


## What percent of the total is this?
```{r}
n_total <- as.numeric(count(mydata))
pctperfectmath <- 100 * (______ /n_total)

```

The percent of the schools with a perfect pass rate on the math test 
is **`r pctperfectmath`** out of a total of **`r n_total`** schools in the sample.


# Part (iii)
## How many have a pass rate of exactly 50%?
```{r}

mydata.exact50math <- filter(mydata,math4==50)
n_exact50math <- as.numeric(count(mydata.exact50math))

## Add your own commands to this r chunk to be able to fill in the blank in the
## sentence below.




```

The number of schools with a perfect pass rate on the math test is
**`r n_exact50math`**, which is **`r ___ `** percent of the total.

```{r}
## Part (iv)
???  <- mean(mydata$math4)
???  <- mean(mydata$read4)
```

The average pass rate for all Michigan schools is **`r ???`** for Math and
**`r ???`** for reading.

## Part (v) Correlation
```{r}

corr_math_and_read <- cor(mydata$math4,mydata$read4)

```

The correlation between math and reading pass rates is **[(pick one) strong or not strong]**, **`r _______`**.


## Part (vi) Coefficient of Variation

```{r}
avg_exppp <- mean(mydata$exppp)
sd_exppp  <- sd(mydata$exppp)
cov_exppp <- sd_exppp/avg_exppp    # example of using R as a calculator
```

The coefficient of variation in per pupil expenditures is low, **`r ___`**. In other words, there is **[(pick one) a lot or not a lot]** of variation
among the schools in how much they spend per pupil. 


## Part (vii) Log Approximation for % change
```{r}

ApctmorethanB_exppp <- 100 * (6500 - 4000) / (4000)
ApctmorethanB_exppp_logapprox <- 100 * (log(6500) - log(4000))

```

[NOTE: The following sentence illustrates the "\frac{}{}" syntax to make
a fraction show up in a sentence.]

Using the percent difference formual $100 * \frac{A - B}{B}$ yields **`r ApctmorethanB_exppp`**. 

Using the difference in logs approximation
yields **`r ApctmorethanB_exppp_logapprox`**.

It matters how the differences in school expenditures are calculated. 
