---
title: "Econ 343 Ch 4 Report Lastname, firstname"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```

Instructions: 

- In addition to using the results for the homework on Cengage, 
you will write up the analysis for both computer exercise assignments (C1 and C8) as a "report" using this template.
- Write of the results for each part of the homework in a sentence or two.
- **Show all work by doing
math in R chunks and using inline R to refer to the result in the write-up.** 

# Ch 4 C1 - Voting Report

## 1. Economic Question

*put the economic question for this problem in your own words*

## 2. Economic Model

*{Instructions: Replace depvar and variables with actual names.}*

$$depvar = f(variables)$$

## 3. Econometric Model

*{Instructions: Replace depvar and x_ vars with the actual variable names.}*

$$depvar = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + u $$

## 4. Data

*Describe the source of the data* 



## 5. Results


### Part (i)


```{r}
# Download the dataset from Cengage and then upload the dataset to Rstudio
# change the location and version number of the dataset 
# in the load command, as needed

load("~/ECON 343/Homework/Data/vote1_v2.RData")
mydata <- Data
stargazer(as.data.frame(mydata), type = "text")  

```



```{r}
# Enter the R commands you need to answer this part of the question
# Do this as needed for all of the subsequent parts

lm1 <- lm(voteA ~ log(expendA) + log(expendB) + prtystrA, data = mydata)
bhat1 <- coef(lm1)

stargazer(lm1,type = "text")

```
$$\widehat{depvar} = 
  `r round(bhat1["(Intercept)"],2)`
+ `r round(bhat1["log(expendB)"],2)`log(expendB) 
+ `r round(bhat1["log(expendA)"],2)`log(expendA) 
+ `r round(bhat1["prtystrA"],2)`prtystrA $$

### Part (ii)

``` {r}
# Compute results as needed and save.
# Refer to the variable name that contains the computed results in
# the inline r chunk in the write up below.

```



### Part (iii)

``` {r}
# Compute results as needed and save.
# Refer to the variable name that contains the computed results in
# the inline r chunk in the write up below.

```

### Part (iv)

``` {r}
# The Cengage problem asks you to work through the algebraic approach to test 
# a linear combination of parameters. 

# Step 1: create the variable dlexp
# Step 2: replace log(expendB) with dlexp and re-run the regression
# Step 3: use the estimated coefficient on log(expendA) to test the hypothesis

mydata <- mydata %>%  mutate(dlexp = lexpendA - lexpendB)
```



---


# Ch 4 C8 - Pensions Report


## 1. Economic Question

*write your answer here*

## 2. Economic Model

$$depvar = f(variables)$$

## 3. Econometric Model
*{Instructions: Replace depvar and x_ vars with the actual variable names.}*

$$depvar = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + u $$

## 4. Data

*Describe the source of the data* 


## 5. Results


```{r}
# Download the R dataset from Cengage and then upload the dataset to Rstudio
# change the location and version number of the dataset 
# in the load command, as needed

load("~/ECON 343/Homework/Data/401ksubs_v1.RData")

# Notice in Cengage that the problem specifies only to use single people (fsize = 1)
# The filter command below is how to choose a subset of the data such that fsize=1

mydata <- Data %>% 
  filter(fsize == 1)
stargazer(as.data.frame(mydata), type = "text")  

```


### Part (i)
```{r}
# Compute results as needed and save.
# Refer to the variable name that contains the computed results in
# the inline r chunk in the write up below.

num.single <- nrow(mydata)

# Use the following in an inline r chunk to include a comma separater
# formatC(as.numeric(num.unmarried),big.mark = ",")
# the "as.numeric" command is needed sometimes

```
The number of single people in the dataset is 
**`r format(num.single,big.mark = ",")`**.


### Part (ii)

``` {r}
# Note: you need to adjust the following regression command to only use single people

lm2 <- lm(nettfa ~ inc + age, data = mydata)
bhat2 <- coef(lm2)

stargazer(lm2,type = "text")



```

**The
estimated equation for part (ii) is:**

$$\widehat{depvar} = 
  `r round(bhat2["(Intercept)"],2)`
+ `r round(bhat2["inc"],2)`inc 
+ `r round(bhat2["age"],2)`age $$

### Part (iii)

``` {r}
# Compute results as needed and save.
# Refer to the variable name that contains the computed results in
# the inline r chunk in the write up below.

```

### Part (iv)

``` {r}
# Compute results as needed and save.
# Refer to the variable name that contains the computed results in
# the inline r chunk in the write up below.


```


### Part (v)

``` {r}
# Compute results as needed and save.
# Refer to the variable name that contains the computed results in
# the inline r chunk in the write up below.

```
