---
title: "Econ 343 Visualize pvalues"
author: "B. Haney"
date: "9/20/2020"
output:
  pdf_document: 
    dev:  png
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearHypothesis(), predict() hccm() heterosced and qqPlot()
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
options(scipen=999) # prevents scientific notation
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```
# Use Visualize.it() to graph p-values

## Requires the package "visualize"

## Note the change to the yaml code to add `dev:  png`

- Use this to visualize the p-values or critical values for hypothesis tests.

- Provide the test-statistic (tstat) and the other information 
required for the relevant test distribution (normal, t, F, $chi^2$, etc.).

- See the specific Visualize command for each distribution.

- Use the "section=" option to fill in either the upper tail,
lower tail, or both:
  - section="upper"
  - section="lower"
  - section="tails"

\newpage

# Visualize.norm()

When using the "tails" option, the tstat and -1*tstat must be
provided as a list "c(-1*tstat,tstat)" as in the following example.
```{r}
tstat = 1.75
visualize.norm(stat=c(-1*tstat,tstat), mu=0, sd=1,section="tails")
```
\newpage
# Visualize.t() 

```{r}
tstat = -1.75
df = 30
visualize.t(stat=tstat,df=df,section="lower")
```
\newpage
# Visualize.F()

```{r}
tstat = 1.75
df1 = 3
df2 = 30
visualize.f(stat=tstat,df1=df1,df2=df2,section="upper")
```

