---
title: "Econ 343 Ch 7 HW C4 Lastname, firstname"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(sjmisc)
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(scipen=999)
options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```

**Instructions:**

- The primary learning objective for this HW is interpreting dummy variables.
- Save your commands and show all work for the Cengage computer exercises
- the variable names should make it obvious what you are computing

- Each problem should begin by providing the context - that is, the steps of empirical analysis:

1. Economic Question
2. Economic Model
3. Econometric Model
4. Data
5. Analysis

- Then begin inserting the parts of the Cengage problem
- perform all calcuations in R chunks
- present all results in the write-up below the r chunks - using r inline to recall any estimated or calculated results

- For each part in the problem:
  - Include the following sections as needed:


_Description_ Explain what the purpose of that part of the question is

_Estimated Equation_ FOR THIS ASSIGNMENT - YOU DO NOT NEED TO WRITE OUT the estimated regression equation using R inline. BUT, always display the regression results using stargazer(). Display regression results side-by-side when they are being discussed together, for example, in part (iii) display the regression from part (ii) and part (iii) using stargazer().

_Results_ 
    - Interpret what the estimated coefficient means, put into real terms when possible and/or helpful
    - Present the results and conclusion of any hypothesis tests

# Ch. 7 C4 Athletes and GPA

## Part (i)

(Note: No R coding needed. Just explain which answer you chose to the question on Cengage and why.)

## Part (ii)
``` {r}

#Hint: use the load() command to load the data from Cengage.

# run the regression requested and display the results using stargazer()

```
(Reminder: You do not need to write out the full estimated equation in R inline for any parts of this assignment.)

(NOTE: For the interpretation section, you only need to interpret the coefficient on *athlete*.)

## Part (iii)
``` {r}
# run the regression requested and display the results using stargazer()


```

(NOTE: For the interpretation section you only need to describe and interpret what happens to the coefficient on *female* when *sat* is removed from the regression. ** 


## Part (iv)
``` {r}

# Hint use mutate with "if_else" (or case_when) to create the femath, maleath, and malenoath variables



```

(Note: In the interpretation section, you only need to provide an interpretation for the estimated coefficient on *malenoath*.)


## Part (v)

``` {r}
# Hint: use mutate to create the femsat variable



```

(Note: In the interpretation section, you only need to discusss hypothesis related to the coefficient on *femsat*)

