---
title: "Econ 343 Midterm 1 Practice Exam DEMO"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```
# Instructions

This practice exam uses the CENGAGE Ch 03 C4 assignment questions for practice.

- Follow the directions in each of the following sections. 
- Look for "==>" and "???" to indicate places you need to supply commands or
add interpretation
- Knit to .pdf and upload to Moodle.


## Create *mydata* using the *attend* dataset from the wooldridge package.

```{r}


```


## Display the descriptive statistics 

- use stargazer with the *mydata* dataset

```{r}


```

### Describe the dataset and the variables


## Economic Question

### How correlated is class attendance with the final exam grade, controlling for the cumulative GPA prior to the term?

## Economic Model

### ??? 

## Econometric Model

$$  ??? $$

### Model 1 Regression Results

- ==> Run the regression and display the results using stargazer
- ==> Round all numbers in the results table to 2 digits
- ==> Save the estimated coefficients from the regression

```{r}

# lm1 <- ???


# bhat1   <- ???
# bhat1_0  <- ???
# bhat1_atnd <- ???
# bhat1_pgpa <- ???

```

### Estimated Regression Equation

- ==> Round all values to two decimal places.

$$  ???  $$


## Interpret the coefficient on attendance rate


INTERPRETATION: ???


## Model 2 - Estimate the regression without controlling for prior GPA

### Model 2 Regression Results

- ==> Run the regression and display the results of both regressions side-by-side
- ==> Round all numbers in the results table to 2 digits

```{r}



```

## Explain the difference in the results

???

Model 2 is **[ ==> pick one: overespecified or underspecified]** because it is missing a key control variable.

The estimated coefficient on attendance rate in Model 2 is biased
**[ ==> pick one: upward or downward]** 

because:

- the expected correlation between GPA and attendence is 
**[ ==> pick one: positive or negative]**

and 

- the expected correlation between GPA and final exam score is 
**[ ==> pick one: positive or negative]**.


