---
title: "Econ 343 Empirical Project lastname, firstname"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_collapsed: yes
    toc_float: yes
    toc_levels: 2
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}

#
# Always include these packages in R projects for this class
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.

library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(haven)
library(expss)      # apply_labels() cro()

library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values

library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data


options(digits=4)   # limits number of decimals to 4 when printing

knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in
options(knitr.duplicate.label = 'allow')

```

UPDATE: 7.11.2020 - revised to use example from GSS data
UPDATE: 7.11.2020 - revised to read in saved file from ECON 343 Exploratary Analysis.Rmd

# Project Summary

{Instructions for Project Summary: Fill in the bold aspects of the 5 parts of empirical analysis below. For each bullet point, provide a sentence or two of description or explanation.}

*1. Economic Question*

**What type of workers tend to work more often in teams?**

- Explain briefly why this question is interesting using infographic, statistics, and opinion articles. 

*2. Economic Model*

**$$\textbf{teamwork} = f(\textbf{education,age,sex,region,college grad})$$**

- Use the research paper you found to help explain the list of explanatory variables.

*3. Econometric Model*

**$$depvar = \beta_0 + \beta_1 xvar_1 + \beta_2 xvar_2 + ... + u $$**

- describe your econometric model in words 

*4. Data*

 **summary tables and crosstabs** for each variable
 
- Define each of the variables.

*5. Results*

**table of final regression results**

- Describe the main findings


# Introduction
{Instructions: For this and all subsequent sections, follow the directions and use the question prompts to help guide what needs to be written up. The example dataset and commands that are provided in the R chunks below are to demonstrate the types of commands you will need to use with your data. The examples should be deleted in the final report.}


- In this section of the project, describe the **economic question** you
are exploring. Explain why this 
question is interesting using statistics, graphs, and the variety of opinions you found.

- Present the **economic model**. What is the dependent variable of interest? 
What are the explanatory variables in the model?


- Briefly summarize what is known so far based on your literature review. 
Make sure you add every citation to your references at the end of the report.

- Describe in detail how your primary paper contributes to understanding this
question. What does it conclude? What data does it use? 

- Evaluate the paper. How robust are its conclusions? What are the strengths 
and limitations of the data and methods. 

- Set up the motivation
for how you modify in the model for your second set of results.

- Briefly explain how your new model differs from the first model. 

- Briefly summarize the result of your comparison.

# Data

Fully describe the data that you are using. What are the questions
that created the variables? What is the time frame? Who are the 
respondents? 

If using the GSS, make sure to cite the data source using a footnote.\footnote{Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen, General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese, and Stephen Morgan; Sponsored by National Science Foundation. --NORC ed.-- Chicago: NORC, 2018: NORC at the University of Chicago [producer and distributor]. Data accessed from the GSS Data Explorer website at gssdataexplorer.norc.org.}

## Description
- What are these data? Where were they obtained? How were the data  
sampled (the "sampling frame"). 

- Define the primary dependent variable(s) and explanatory variables. 

- Explain how these variables are operationalized in the dataset.

- Make sure to run your **Empirical Project exploratory analysis.Rmd** file to create 
and save your ***mygss*** dataset before loading it below.

```{r}

# Display the basic variables you are using for your analysis 
load("~/ECON 343/Empirical Project/mygss.RData")
mydata <- mygss
stargazer(mydata,type = "text")
as_tibble(mydata)
```

# Preliminary Analysis

What are the key relationships you want to explore? Look at simple comparisons and correlations first.

**What types of workers tend to work in teams more than independently?**

```{r}
## Use this command as a template to produce cross-tabs for Qualitative variables

mydata %>%
  calc_cro_cpct_responses(list(teamwork),              # list a qualitative dependent variable
                          list(SEX,total())) %>%     # list a qualitative explanatory variable
  significance_cpct()   # test for significant differences

mydata %>%
  calc_cro_cpct_responses(list(teamwork),              # list a qualitative dependent variable
                          list(agegrp,total())) %>%     # list a qualitative explanatory variable
  significance_cpct() 
```

**Discuss** There are statistically significant differences by age (workers over age 65 are 
less likely to be in team-work oriented jobs), but no differences between men and women.


**Economic Question: Does statistical significance patterns change when controlling for agegrp or gender?**
```{r}
mydata %>%
  calc_cro_cpct_responses(list(teamwork),   # qualitative dependent variable
                          list(agegrp %nest% SEX,total())) %>%     # add a control variable
  significance_cpct()  

mydata %>%
  calc_cro_cpct_responses(list(teamwork),   # qualitative dependent variable
                          list(SEX %nest% agegrp,total())) %>%     # add a control variable
  significance_cpct()  

```

**Discuss** The difference between men and women is still not statistically signficant
when controlling for age. Age is still stat. sig. even after controlling for male vs. female.
  
**Economic question: What other attributes of a worker appear to be associated with having a teamwork-oriented job?**

```{r}
# Use this command as a template to compare Means of Quantitative variables by categories
mydata %>%
  tab_cols(teamwork) %>%                        # columns represent categories to compare
  tab_cells(INCOME,WEALTH,AGE,EDUC) %>%            # cells are the continuous variables
  tab_stat_mean_sd_n() %>%
  tab_last_sig_means() %>%                    # test for significant differences 
  tab_pivot()
```

**Discussion** Age appears to be the only relevant variable. However, the statistical
significance of the other variables might show up if age is controlled for. Let's see:


```{r}
mydata %>%
  tab_cols(agegrp %nest% teamwork)     %>%            # "nested" columns are categories within categories
  tab_cells(INCOME,WEALTH,AGE,EDUC)    %>%            # cells are the continuous variables            
  tab_stat_mean_sd_n()            %>%
  tab_last_sig_means()            %>%
  tab_pivot()

```

**Discussion** Higher levels of education do seem to counteract the age effect, a little. 
The only group that is statistically significantly less likely to have a teamwork-oriented
job are workers over age 65. However, within that group, more highly educated workers
are more likely to have a teamwork-oriented job than less-educated over 65 year olds.

Now it is time to use regression analysis to control for several attributes at once and
estimate the marginal effects of each of these factors on having a teamwork-oriented job.

# Regression Analysis

## Dummy variables

  - For regressions, collapse multiple categories down to two (0/1)
  - use the `val_lab()` command to find out what you want to recode
  - Use `as.numeric()` with variables that have labels as their values


```{r}

# Here is an example for how to collapse multi-categorical dependent variables
# into a dummy variable to run a Linear Probability Model

val_lab(mydata$COURTS)
val_lab(mydata$DEGREE)
val_lab(mydata$region)

mydata <- mydata %>%
  mutate(not_harsh_enough = case_when(COURTS == 1 | COURTS == 3 ~ 0,
                                      COURTS == 2  ~ 1),
         collgrad = case_when(DEGREE == 3 | DEGREE == 4 ~ 1,
                                DEGREE == 1 | 
                                DEGREE == 2 | 
                                DEGREE == 3 ~ 0),
         south = case_when(region == 3 ~ 1,
                             region == 1 | 
                             region == 2 | 
                             region == 4 ~ 0),
         female = case_when(SEX == 2 ~ 1,
                            SEX == 1 ~ 0))

cro(mydata$DEGREE,mydata$collgrad)
cro(mydata$SEX,mydata$female)





```

## Missings

When dealing with real data (outside of homework-land), 
missing value patterns can cause problems. 

- Regressions only use observations that
have no missings in any of the variables in the regression. 

- Post-regression commands, such as fitted(), want to use the
entire dataset, includig observations that are not in the regression.

- So, always 
create a fresh dataset for your regression that includes only observations with no missing values for the variables
in your model. 

- See the `na.omit() with the select()` command below to create `lm.mydata` that contains only observations with no missing values for the variables used in regressions 

- Use `lm.mydata` for all regressions to ensure that the the results are based on the same sample 

- Use `lm.mydata` for post-regression analysis to ensure you are using the same dataset consistently


```{r}
lm.mydata <- mydata %>% select(teamwork,EDUC,female,AGE,agegrp,collgrad,INCOME,region)
lm.mydata <- na.omit(lm.mydata)

stargazer(lm.mydata,type = "text")
# as_tibble(lm.mydata)
```


## Regression Results
```{r}
lm1 <- lm(teamwork ~ EDUC + AGE + female + collgrad, data = lm.mydata)

stargazer(lm1,type="text")

```


## Check Predicted probabilities

Remember to use the lm.mydata data in all regressions, and post-regression commands
```{r}
# Get the "y-hats"
# Recode y-hats to be dummy y_hat (take on 0/1 values only)

lm1.data <- lm.mydata %>%
  mutate(yhat  = fitted(lm1),
         yhat_dummy = case_when(yhat >= .5 ~ 1,
                                yhat  < .5 ~ 0))

range_yhat <- range(lm1.data$yhat)
gt1_yhat   <- count(lm1.data$yhat > 1)

cro_cpct(lm1.data$teamwork, lm1.data$yhat_dummy)

```

**Write-up** The fitted probabilities range from `r range_yhat[1]` to `r range_yhat[2]`, so none are negative.  
There are `r gt1_yhat` predicted probabilities above 1..


# Evaluate the LPM model 

**How well does this model predict team-oriented job?**

Compare actual to predicted probabilities.

```{r}
# create a variable that determines if the predicted y_hat was correct or not
lm1.data <- lm1.data %>%
  mutate(correct_pred = case_when(yhat_dummy == teamwork ~ 1,
                                  yhat_dummy != teamwork ~ 0))

# overall correct prediction rate
Pct_correct_pred <- mean(lm1.data$correct_pred)

# percent of ecobuyers correctly predicted
Pct_correct_teamworker <- mean(lm1.data$correct_pred[lm1.data$teamwork == 1]) * 100

# percent of non-ecobuyers correctly predicted
Pct_correct_soloworker <- mean(lm1.data$correct_pred[lm1.data$teamwork == 0]) * 100
                            
```

**Write-up** This model does a decent job at predicting beliefs about court sentencing. But, much better for predicting those who believe sentences should be harsher than those who don't.

Using the standard prediction rule – $\hat{y}_{dummy}$ = 1 if $\hat{y}$ > or = .5 and zero otherwise:

- The overall percent correctly predicted is **`r Pct_correct_pred`%.**

- The percent correctly predicted for team workers is
**`r Pct_correct_teamworker`%**.   

- The fraction correctly predicted for solo workers is 
only **`r Pct_correct_soloworker`**%.  

# Sensitivity Analysis

- Try running your model on a sub-group or re-code a variable differently to see if you still get 
qualitatively similary results

- Compare and contrast the results of the two approaches. Are your results "robust" to an alternative model?

For this example, let's see if the regression is different if we run exclude workers living in the West (silicon valley).

```{r}
cro(lm.mydata$region)

lm2 <- lm(teamwork ~ EDUC + AGE + female + collgrad, 
          data = lm.mydata %>% 
            filter(region != "West"))

stargazer(lm1,lm2,type="text")

```

**Discuss** Removing the workers in the "West" did not change the results by much at all.

# Linear Hypothesis Test

- Test something from your main model or use an F-test to compare the two models you ran for sensitivity analysis.

Example: Neigher education or the dummy variable for college graduate are statistically
significant in the models. But, might they be jointly statistically significant?

```{r}
h1 <- linearHypothesis(lm1,c("EDUC=0","collgrad=0"))
h2 <- linearHypothesis(lm2,c("EDUC=0","collgrad=0"))
h1
h2

```

**Discuss** Education and college-degree are neither individually nor jointly statistically signficant. There is not enough evidence to suggest that education is related to
whether a person works in a team or solo on their job.

# Conclusion

- In this last section:
  + Restate the economic question
  + Explain how the primary research paper contributed to understanding that
  question.
  + Describe what you learned from the process
  + Conclude with suggestions of additional questions that would be
  interesting to pursue based on what you found.
  
# References
