---
title: "Empirical Project Part 1 ** Change this!! Lastname Firstinitial **"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_collapsed: yes
    toc_float: yes
    toc_levels: 4
---

```{r setup, include=FALSE}
library(wooldridge) # datasets in the textbook
library(dplyr)      # data wrangling
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(expss)      # apply_labels() calc_cro() tables
library(sjmisc)     # frq()
library(sjlabelled) # useful label commands
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data


options(digits=4)   # limits number of decimals to 4 when printing

knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```

# Get ready! 

When starting with real data, the first step is to get the data cleaned and organized. You will handle missing values and create new
variables and categories based on what you need. Sometimes you will
combine data from different datasets together. All of this is 
known as "data wrangling".

## Read in the dataset

The following commands read in the entire GSS 2018 dataset that has been download for you already. Your task is to add additional variables that you might want to analyze to the list of variables in `myvars`.

```{r}
## 
## >>>> STUDENT WORK: Dont' forget - KNIT to HTML (not.pdf)
##

## 
## >>>> STUDENT WORK: Update the location of the dataset as necessary:
##
load("~/ECON 343/Empirical Project/GSS2018.Rdata")

## Don't change the variables in "varsstd" - everyone uses these
varsstd <- c("YEAR", "ID", "AGE","RACE","SEX","MARITAL","CHILDS",
             "REGION","EDUC","DEGREE","WRKSTAT","REALINC","REALRINC",
             "PARTYID","TAX","COURTS",
             "WTSSALL")
## 
## You add additional variables to the "myvars" list below as needed
##

myvars    <- c( )  ##<<<< fill in the variables you want here 

varlist <- c(varsstd,myvars)

mygss <- gss2018 %>% select(all_of(varlist))
 
 
stargazer(mygss,type="text",
          digits = 0, 
          summary.stat = c("n","min","max","median","mean","sd"),
          title = "MYGSS")
```



## Recode missings, Collapse categories, and create new variables"

The first data wrangling task is to recode missings.

- You can examine the value labels of the variables to identify actual values that
should be recoded as missing. For example, '9' is a common value that denotes "Refused to answer." You want it to be recoded to "NA" (R's term for a missing value)

- You can also refer to the codebook on the GSS website to obtain even more
detailed information and the actual survey questions for the variables:
https://gss.norc.org/Get-Documentation

- Look for values that will need to be recoded as missing (e.g. "NA", "IAP" (for inapplicable), "DK (for don't know)", R doesn't have to pay", etc.)
- Look for categorical variables you will want to collapse to fewer categories
- Look for measurement variables that you might want to create categorical variables from

## Print out Value Labels

First look at the categorical variables.
```{r type="latex"}

## Add any or your own new categorical variables to the list below

catvars  <- c("RACE", "SEX","DEGREE","WRKSTAT") ##<<< add your categorical variables here

for (x in catvars) {
  table <-  frq(mygss[[x]],
                weights =mygss$WTSSALL,
                title=paste(x,var_lab(mygss[[x]])))
  print(table)
}

```


## Create new variables 

- You will often want to create new variables so that you can
reduce the number of categories of a variable. 
- Be sure to
document what the categories mean so that you can
label them correctly.
- Use `mutate()` with `case_when()` to collapse several categories into a few

`Usage:

datasetname <- datasetname %>%
     mutate(**newvarname** = 
     case_when(condition 1 ~ value if condition 1 holds, 
     condition * ~ value if condition 2 holds))`


The following code chunk creates new variables *party*, *region*, *labor force status*, and *agegrp* by collapsing the values of PARTYID, REGION, WRKSTAT, and AGE into a smaller number of categories, using the category cutoffs in the GSS Trend Graphs.

_NOTES_

- use labels for the recoded values, rather than numbers
- use 0/1 (not characters) when creating dummy variables

``` {r}

# The GSS variable PARTYID was recoded 
# into four categories: (0)-(1) Democrat; (2)-(4),(7) Independent, Other; and (5)-(6) 
# Republican. 

# The GSS variable REGION was recoded into four categories: (1)-(2) Northeast; 
# (3)-(4) Midwest; 
# (5)-(7) South; (8)-(9) West. These are in alignment with the Census regional groupings.

# The GSS variable WRKSTAT was recoded into three categories: 
# (1)-(3) Employed, (4) Unemployed, 
# and (5)-(8) Not in labor force/Other. 
# This is in alignment with the Bureau of Labor Statistic's labor force definitions.

mygss <- mygss %>%
  mutate(party = case_when(PARTYID == 5 | PARTYID == 6 ~ "Republican",      
                           PARTYID == 0 | PARTYID == 1 ~ "Democrat",      
                           PARTYID >= 2 & PARTYID <= 4 ~ "Independent",      
                           PARTYID == 7 ~ "Independent"),
         
         region = case_when(REGION == 1 | REGION == 2 ~ "NE",     
                            REGION == 3 | REGION == 4 ~ "Midwest",    
                            REGION == 5 | REGION == 6 | REGION == 7 ~ "South", 
                            REGION == 8 | REGION == 9 ~ "West"),    
         
         # LFS = Labor Force Status
         lfs    = case_when(WRKSTAT <= 3 ~ "Employed",                    
                            WRKSTAT == 4 ~ "Unemployed",                    
                            WRKSTAT > 4 & WRKSTAT < 9 ~ "OLF"),      

         agegrp = case_when(AGE >= 18 & AGE <= 34 ~ "Age 18-34",
                            AGE >= 35 & AGE <= 49 ~ "Age 35-49" ,
                            AGE >= 50 & AGE <= 64 ~ "Age 50-64",
                            AGE >= 65 ~ "Age 65+"))

# Check accuracy of recodes
cro(mygss$PARTYID,mygss$party)
cro(mygss$REGION,mygss$region)
cro(mygss$WRKSTAT,mygss$lfs)

```



**Dummy variables**

Dummy variables are "0/1" variables, with value = 1 if the variable is true
and value = 0 if the variable is false

- give dummy variables values of 0/1 rather than categorical names so that you have better control over which value of the variable is in the regrssion.

``` {r}

                              
mygss <- mygss %>%         
         mutate(married  = case_when(MARITAL == 1 ~ 1,
                              MARITAL  > 1 ~ 0),
          haschild = case_when(CHILDS   > 0 ~ 1, 
                               CHILDS == 0 ~ 0),
         
         tax.toohigh = case_when(TAX == 1 ~ 1,
                                 TAX == 2 | TAX == 3 ~ 0))

cro(mygss$MARITAL,mygss$married)
cro(mygss$CHILDS,mygss$haschild)
cro(mygss$TAX,mygss$tax.toohigh)
```


## Label variables
Variable labels are important for making tables and results "human readable". 

Use `apply_labels()` to:

- Name or rename variables

`Usage:

- datasetname <- **apply_labels**(datasetname,*var1* = "var label 1")`

``` {r}
mygss <-  apply_labels(mygss,
    party  = "Political Party",
    region = "Region",
    lfs    = "Labor Force Status",
    agegrp     = "Age",
    married    = "Married",
    haschild   = "Has Children",
    tax.toohigh = "Taxes are too darn high!"
  )
```

## Inflation adjust the income variable

- `REALINC` is total family income 
- `REALRINC` is respondent's income
- `REALINC` and `REALRINC` are in 1986 USD. 

- to convert to 2018 USD, multiply by the CPI factor of 2.26 (see: https://www.bls.gov/data/inflation_calculator.htm) 

```{r}
mygss <- mygss %>%
  mutate(faminc2018 = 2.26 * REALINC,
         respinc2018 = 2.26 * REALRINC)

stargazer(mygss %>% select(REALINC, faminc2018, REALRINC, respinc2018),type = "text")

gf_histogram(~ faminc2018, data = mygss)
gf_histogram(~ respinc2018, data = mygss)

```

# Get set!

Take stock of the variables you have to work with. 

Use the following `cro()` commands from the `expss` package to look at the frequencies of categorical variables.

- `cro()` with one variable displays the for each level of a factor variable
- `cro_cpct()` reports percentage responses for each level of a factor variable
- `cro_cpct(var1,var2)` with two variables. The second variable is the explanatory variable
- use `cro_cpct()` with `%nest%` for more than one explanatory variable (see below)
- use `drop_rc()` in front of all `cro()` and related table commands to prevent the display of empty rows or columns

## Dependent Variables

Add yours here:

``` {r}

drop_rc(cro_cpct(mygss$tax.toohigh, weight = mygss$WTSSALL))

```


## Demographic Variables

Add yours here: 

``` {r}
drop_rc(cro(mygss$SEX, weight = mygss$WTSSALL))
drop_rc(cro_cpct(mygss$RACE, weight = mygss$WTSSALL))
drop_rc(cro_cpct(mygss$agegrp, weight = mygss$WTSSALL))
drop_rc(cro_cpct(mygss$region, weight = mygss$WTSSALL))
drop_rc(cro_cpct(mygss$married, weight = mygss$WTSSALL))
drop_rc(cro_cpct(mygss$haschild, weight = mygss$WTSSALL))

```

##  Explanatory Variables

Add yours here:
``` {r}

drop_rc(cro_cpct(mygss$DEGREE, weight = mygss$WTSSALL))
drop_rc(cro_cpct(mygss$lfs, weight = mygss$WTSSALL))
drop_rc(cro_cpct(mygss$party, weight = mygss$WTSSALL))


```

## Graphs

```{r}
gf_bar(~ party, data = mygss)
```


# Go! 

Explore the variables to:

- Look for statistically signficiant differences across groups
- Look for particularly important control variables

## Add statistical significance to R tables:

- use `cro_cpct()` with two variables
- add the pipe command `%>%`
- add the `significance_cpct()` option 
to display statistical signficance between columns

How to read the table:

- the table columns are labeled with letters. 
- the letters next to a value in the table indicates that the value of the variable for this group is statistically significantly larger than the value for the group in a different column.

## Age and Political Affiliation example
``` {r}

drop_rc(cro_cpct(mygss$agegrp,mygss$party, weight = mygss$WTSSALL))%>%
  significance_cpct()

```

## COURTS example

Suppose you want to see if
beliefs about *COURTS* appear to be related to the *region* of the
respondent
``` {r}
drop_rc(cro_cpct(mygss$COURTS,mygss$region, weight = mygss$WTSSALL)) %>% 
  significance_cpct()

```

It appears that the percent of respondents in the Midwest that believe courts are 
"about right" is statistically significantly lower than all other regions. The
percent in the 
Midwest that believe courts are "not harsh enough" is higher than all other regions,
but that difference is not statistically significant.


## %nest% command

Are the differences statistically significant after controlling for race?

Use the %nest% command to compare categories within another category

Usage:

- cro(*row variable*, *top level* %nest% *bottom level* )

``` {r}
drop_rc(cro_cpct(mygss$COURTS, mygss$RACE %nest%  mygss$region, weight = mygss$WTSSALL)) %>%   
  significance_cpct()

```

After controlling for race, it appears that the percentage of whites in the Midwest
how believe courts are "not harsh enough" is statistically significantly 
greater than the percent of whites in other regions. What do you notice in these
results?



# Your turn

What relationships do you want to explore? Choose one that you find interesting
and share it on the Community Learning forum this week. You can take a screen shot
of the table, or print the html page as a .pdf and upload it to the post.

You might want to recode some of the variables differently. Or, you might even want
to explore the GSS for additional variables. You can add new variables from the
GSS by adding their name to the "select()" statement when reading in the dataset.

``` {r}



```


# Save the final version of your mygss file so you can use it in your report.

_IMPORTANT_

- Save the file you created during this exploratory
analysis and then read it in at the start of your final empirical project report.

```{r}
save(mygss,file="mygss.RData")

```

