---
title: "Econ 343 Empirical Project Final Report **lastname, firstname**"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
    toc_collapsed: yes
    toc_float: yes
    toc_levels: 2
---


```{r setup, include=FALSE}
library(wooldridge) # datasets in the textbook
library(dplyr)      # data wrangling
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(expss)      # apply_labels() calc_cro() tables
library(sjmisc)     # frq()
library(sjlabelled) # useful label commands
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data
library(kableExtra)

options(scipen=999)
options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunk echoed in
options(knitr.duplicate.label = 'allow')
```


# Project Overview

{Instructions for Project Summary: Fill in the bold aspects of the 5 parts of empirical analysis below. For each bullet point, provide a sentence or two of description or explanation.}

## 1. Economic Question*



- Explain briefly why this question is interesting using infographic, statistics, and opinion articles. 

## 2. Economic Model


- Helpful Hint: Use the research paper you found to help explain the list of explanatory variables.

## 3. Econometric Model

- describe your econometric model in equation and words 

## 4. Data


- describe the GSS and the year you are using for your analysis

- make sure to cite the GSS data source using a footnote.\footnote{Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen, General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese, and Stephen Morgan; Sponsored by National Science Foundation. --NORC ed.-- Chicago: NORC, 2018: NORC at the University of Chicago [producer and distributor]. Data accessed from the GSS Data Explorer website at gssdataexplorer.norc.org.}

- describe who was sampled in the dataset

- describe how the data are weighted

- define each of the variables that you use in your analysis.

- provide **summary tables and crosstabs** for each variable

- include important tables from Part 1 that demonstrate relationships between variables that you will explore further with regression analysis
 

## 5. Summary of Results

- this is like an abstract

- in a paragraph or so describe what you did in your analysis and display a table or graph of one of your key findings

- the next part of the empirical report contains the full methods, analysis, and conclusion sections

# Introduction

- In this section of the project, go into more detail about the **economic question** you
are exploring. 

- Explain why this 
question is interesting using statistics, graphs, and the variety of opinions you found.

- Present the **economic model**. What is the dependent variable of interest? 
What are the explanatory variables in the model?

- Briefly summarize how the research paper(s) you found contribute to understanding your question. 
Make sure you add every citation to your references at the end of the report.



## Data and Methods
- What are these data? Where were they obtained? How were the data  
sampled (the "sampling frame"). 

- Define the primary dependent variable(s) and explanatory variables. 

- Explain how these variables are operationalized in the dataset.


```{r}
# Display the basic variables you are using for your analysis
# Change the location of your mygss.RData file as needed
load("~/ECON 343/Empirical Project/mygss.RData")
mydata <- mygss
stargazer(mydata,type = "text")


```

# Preliminary Analysis

- What are the key relationships you want to explore? 

- Look at simple comparisons and correlations first.

- Re-create at least 2 key tables you created in the Empirical Project Part 1.Rmd file

- Format, lable, and title the tables to be presentable for a professional final draft

- see the example table below for how to create tables that can be knitted to .pdf (use the "split_table_to_df()" ...)


```{r}
# Example table
 drop_rc(cro_cpct(mygss$agegrp,mygss$party, 
                  weight = mygss$WTSSALL))%>%
  significance_cpct()%>%
  split_table_to_df() %>%
  kable(booktabs = T) 

```


**Discuss** Discuss the points that you are illustrating for each table.

```{r}



```


# Regression Analysis

## Missings

Instructions:

When dealing with real data (outside of homework-land), missing value patterns can cause problems. 

- Regressions only use observations that have no missings in any of the variables in the regression. 

- Post-regression commands, such as fitted(), want to use the entire dataset, includig observations that are not in the regression.

- So, always create a fresh dataset for your regression that includes only observations with no missing values for the variables in your model. 

- See the example below that uses `na.omit() with the select()` command to create `lm.mydata` that contains only observations with no missing values for the variables used in regressions 

- Use `lm.mydata` for all regressions to ensure that the the results are based on the same sample 

- Use `lm.mydata` for post-regression analysis to ensure you are using the same dataset consistently


## Create regression dataset

```{r}
# Example - be sure to replace or add to the list of variables that you need
# Be sure you include the WTSSALL variable in the dataset you use for regressions

lm.mydata <- mydata %>% select(teamwork, AGE,faminc2018,EDUC,WTSSALL)

lm.mydata <- na.omit(lm.mydata)

stargazer(lm.mydata,type = "text")

```

## Regression Results
```{r}

# Example regression - replace this regression model with your own.
# Be sure to use the command "weights = WTSSALL" in your regression

lm1 <- lm(teamwork ~ EDUC + AGE + faminc2018, 
          data = lm.mydata,
          weights = WTSSALL)

stargazer(lm1,type="text")

```


## Check Predicted probabilities

- Remember to use the lm.mydata data in all regressions, and post-regression commands

- Remember to use the "weight = WTSSALL" option in all tables.

```{r}



```

**Write-up** Describe whether the model produced invalid predicted probabilities or not


# Evaluate the LPM model 


- Compare actual to predicted probabilities.

```{r}




```

**Write-up** Discuss how well the model did at predicting the correct outcomes: overall, and for each of the two different groups of your dependent variable  

# Sensitivity Analysis

- Try running your model on a sub-group or re-code a variable differently to see if you still get 
qualitatively similary results

- Compare and contrast the results of the two approaches. Are your results "robust" to an alternative model?


```{r}



```

**Discuss** Explain how similar your conclusions are under different approaches


# Linear Hypothesis Test

- Test something from your main model or use an F-test to compare the two models you ran for sensitivity analysis.



```{r}


```

**Discuss**



# Conclusion

- In this last section:
  + Restate the economic question
  + Explain how the primary research paper contributed to understanding that
  question.
  + Describe what you learned from the process
  + Conclude with suggestions of additional questions that would be
  interesting to pursue based on what you found.
  
# References
