---
title: "Econ 343 Ch  7 R Lecture"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_collapsed: yes
    toc_float: yes
    toc_levels: 3
  beamer_presentation: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```

# Ch 7 Part 2 R Lecture

## R commands reviewed in this lecture

- partial effect graphs `plot(effect())`

- mutate() with case_when() and if_else()

- cro() for crosstabs

### Key Concepts in this lecture

- Linear Probability Model

- Selection Bias in Treatement effect models



# Linear Probability Model

- A dummy dependent variable takes on the value of "0" or "1". 

- The predicted values (fitted()) from
the regression represent a probability, somewhere between 0 and 1, of belonging to the category. 

- the fitted() values can be re-coded to be "0" or "1" based on whether the predicted probability is greater than some value, usually 0.05

- Regression coefficients represent the change in percentage points of the probability of belonging to the category. 

- In this sense, interpretation in LPM is similar to that of log(y) dependent variables.



## Eco-friendly apples example


**Economic Question** (From Ch 7 Computer Exercise C13) Does the price difference 
between eco- and non-eco-friendly apples affect whether consumers by eco-friendly apples or not?

1. Frequently, you will start by creating the dummy dependent variable by recoding a quantitative variable. 

```{r}
mydata <- apple
mydata <- mydata %>%
  mutate(ecobuy = case_when(ecolbs == 0 ~ 0,
                            ecolbs  > 0 ~ 1))
stargazer(mydata,type="text")
```

**REMINDER: The mean of a dummy variable is the percent of observations with a value of "1".**

In this sample, **62%** of families bought eco-friendly apples.

2. Run the LPM regression, using the dummy as the dependent variable. Save the results for
post-estimation computations.

```{r}
lm7.13 <- lm(ecobuy ~ ecoprc + regprc + faminc + hhsize + educ + age, data=mydata)
stargazer(lm7.13,type="text")

b <- coef(lm7.13)
```

**Write-up** The primary regressor we want to examine is price of eco- and non-eco apples. The coefficient on ***ecoprc*** (price of eco-friendly apples) 
is **$\hat\beta_{ecoprc}$ = `r b["ecoprc"]`** dollars.  This means that if the price eco-apples is
$0.10 higher than regular apples, then the probability of buying eco-labeled apples *falls* by
**.10 * $\hat\beta_{ecoprc} * 100$ = `r 0.10 * b["ecoprc"] * 100` percentage points**.

The coefficient on ***regprc*** (price of regular apples) 
is **$\hat\beta_{regprc}$ = `r b["regprc"]`** dollars. If ***regprc*** increases by **10 cents**, the probability of buying eco-labeled apples **increases** by **.10 * $\hat\beta_{regprc}$ * 100 = `r 0.10 * b["regprc"] * 100` percentage points**.  


# LPM: Check predicted probabilities

Note: In the previous write-up, we must assume that the predicted probabilities 
are not close to the boundaries of zero and one. The next step checks this assumption.

- predicted $\hat{y}$ are predicted probabilities 
  - one concern researchers have about LPM is that it allows invalid predictions (negative or greater than 1)
  - but, the predictions (*fitted()* values) can be checked
  - if there are no or only a few negative or greater than 1 fitted values there is generally no cause for concern

```{r}
# Get the "y-hats"
# Recode y-hats to be dummy y_hat (take on 0/1 values only)

mydata <- mydata %>%
  mutate(yhat  = fitted(lm7.13),
         yhat_dummy = case_when(yhat >= .5 ~ 1,
                                yhat  < .5 ~ 0))

range_yhat <- range(mydata$yhat)
gt1_yhat   <- count(mydata$yhat > 1)
cro_cpct(mydata$ecobuy, mydata$yhat_dummy)

range_yhat[1]
range_yhat[2]
gt1_yhat


                            
```

**Write-up** The fitted probabilities range from `r range_yhat[1]` to `r range_yhat[2]`, so none are negative.  
There are `r gt1_yhat` predicted probabilities above 1, which is not a source of concern with 660 observations.


# LPM Compare actual to predicted probabilities

**Evaluate the LPM model: How well does this model predict eco-buying behavior?**

```{r}
# create a variable that determines if the predicted y_hat was correct or not
mydata <- mydata %>%
  mutate(correct_pred = case_when(yhat_dummy == ecobuy ~ 1,
                                  yhat_dummy != ecobuy ~ 0))

# overall correct prediction rate
Pct_correct_pred <- mean(mydata$correct_pred)

# percent of ecobuyers correctly predicted
Pct_correct_pred_ecobuy <- mean(mydata$correct_pred[mydata$ecobuy == 1])

# percent of non-ecobuyers correctly predicted
Pct_correct_pred_noecobuy <- mean(mydata$correct_pred[mydata$ecobuy == 0])

Pct_correct_pred
Pct_correct_pred_ecobuy
Pct_correct_pred_noecobuy
                            
```

**Write-up** This model does a decent job at predicting eco-buying behavior. But, much better for predicting those who do buy eco-friendly apples, vs. those who won't.

- Use the standard prediction rule – $\hat{y}_{dummy}$ = 1 if $\hat{y}$ > or = .5 and zero otherwise.

- The overall percent correctly predicted is **`r Pct_correct_pred*100`%.**

- The percent correctly predicted for those who are ecobuyers is
**`r Pct_correct_pred_ecobuy*100`%**.   

- The fraction correctly predicted for non-ecobuyers is **`r Pct_correct_pred_noecobuy*100`**%.  

Thus, using the usual prediction rule, the model predicts which consumers will *buy* eco-labeled apples about twice as accurately as it predicts that consumers will not buy eco-apples. 

# Try it yourself: Women labor force participation example

- Example in text explanation for section 7-5

- uses MROZ dataset

## Recreate regression

```{r}

mydata1 <- mroz

# stargazer(mydata1,type = "text")

lpm1 <- lm(inlf ~ nwifeinc + educ + exper + I(exper^2) +
             age + kidslt6 + kidsge6, data = mydata1)
stargazer(lpm1,type = "text", intercept.bottom = FALSE)

```

## Graph partial effect

- for more information about options for effect plots, see https://cran.r-project.org/web/packages/effects/vignettes/predictor-effects-gallery.pdf

```{r}
## This command graphs partial effects at the mean of all other regressors
plot(effect("educ",lpm1))  

## This is how you pick the values of the other regressors and add titles and limits for the y axis
plot(effect("educ",lpm1,
            xlevels=list(educ=c(0,5,10,15,20)),
            fixed.predictors=list(given.values = c(nwifeinc=50,exper=5,
                             age=30,kidslt6=1,kidsge6=0))),
     main = "Education Partial Effects (hccm)",
     sub = "at nwifeinc=50, exper=5, age=30, kidslt6=1, kidsge6=0",
     axes=list(y=list(lim=c(-0.5,1.0),lab=("Predicted female lfp"))))

```

## LPM: Check predicted probabilities
```{r}
# Get the "y-hats"
# Recode y-hats to be dummy y_hat (take on 0/1 values only)

mydata1 <- mydata1 %>%
  mutate(yhat  = fitted(lpm1),
         yhat_dummy = case_when(yhat >= .5 ~ 1,
                                yhat  < .5 ~ 0))

range_yhat <- range(mydata1$yhat)
gt1_yhat   <- count(mydata1$yhat > 1)
cro_cpct(mydata1$inlf, mydata1$yhat_dummy)

range_yhat[1]
range_yhat[2]
gt1_yhat


                            
```


## LPM: Compare actual to predicted probabilities

```{r}
# create a variable that determines if the predicted y_hat was correct or not
mydata1 <- mydata1 %>%
  mutate(correct_pred = case_when(yhat_dummy == inlf ~ 1,
                                  yhat_dummy != inlf ~ 0))

# overall correct prediction rate
Pct_correct_pred <- mean(mydata1$correct_pred)

# percent of ecobuyers correctly predicted
Pct_correct_pred_inlf <- mean(mydata1$correct_pred[mydata1$inlf == 1])

# percent of non-ecobuyers correctly predicted
Pct_correct_pred_notinlf <- mean(mydata1$correct_pred[mydata1$inlf == 0])

Pct_correct_pred
Pct_correct_pred_inlf
Pct_correct_pred_notinlf
                            
```

# Selection Bias

- dummy variable represents treatment

- non-experimental - treatment is a choice

- outcome is a function of not only receiving treatment, but also other characteristics of the participants

- some characteristics can be observed and controlled for (age, experience, etc.)

- some characteristics are not observable (motivation, access to social networks, etc.)



## Job training example

- example 7.13 in section 7-6a

- you have to rename the dataset before running this next r chunk

- rename "jtrain98.RData.rename" to "jtrain98.RData"

```{r}
## for instructor use only: jtrain98 <- as.data.frame(read_dta("jtrain98.dta"))
## for instructor use only: save(jtrain98,file="jtrain98.RData")

load("jtrain98.RData")
mydata2 <- jtrain98

lm.jtrain0 <- lm(earn98 ~ train,data=mydata2)
lm.jtrainRRA <- lm(earn98 ~ train + earn96 + educ + age + married,data=mydata2) 
lm.jtrainURA <- lm(earn98 ~ train + earn96 + educ + age + married +
                    I(train * (earn96 - mean(earn96))) +
                    I(train * (educ - mean(educ))) +
                    I(train * (age - mean(age))) +
                    I(train * (married - mean(married)))
                   ,data=mydata2)

stargazer(lm.jtrain0,lm.jtrainRRA,type = "text",intercept.bottom = FALSE,
          column.labels = c("Regression 1","Regression 2"),
          model.numbers=FALSE,
          column.sep.width = "1pt")

stargazer(lm.jtrainRRA,lm.jtrainURA,type = "text",intercept.bottom = FALSE,
          column.labels = c("Regression 2","Regression 3"),
          model.numbers=FALSE,
          column.sep.width = "1pt")


```

_Conclusion_ 

**Regression 1** Difference in means

- without controlling for observable characteristics of the participant the estimated effect of job training on earnings is negative!  

- this is likely because job training is not randomly assigned. Participants in job training are likely to have lower previous earnings (*earn96*), less education, older, and single.

**Regression 2** Difference in means controlling for educ,age,earn96,married

- adding in control variables for *educ,age,earn96,married* yields a positive and statistically significant effect of job training (the second regression) because now teh job training treatment effect compares earnings for participants and non-participants who have the same background characteristics (same level of previous earnings, education, etc.)

**Regression 3** Average Treatment Effect

- adding interaction terms between the treatment dummy and the de-meaned control variables yields the average partial effect for treatment (**Average Treatment Effect**)

_Caution_

- while adding control variables help greatly to reduce selection bias, there is likely to be selection bias on **unobservables** as well

- controlling for unobservable selection bias is possible with more advanced methods in Ch 17
