---
title: "Econ 343 Ch 2 R Lecture"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```
# Ch 02 Example

### R Commands demonstrated in this lecture:
- gf_point() for scatter plots
- gf_histogram()
- lm()
- stargazer()
- coef()
- fitted()

```{r}
mydata <- wage1  
as_tibble(mydata)  
```

## Explore the data
```{r}
summary(mydata)
stargazer(mydata,type="text")
gf_histogram(~wage,data=mydata)
gf_histogram(~educ,data=mydata)
gf_point(wage ~ educ,data=mydata)
cor(mydata$wage,mydata$educ)

```


## Basic Regression output
```{r}
# OLS regression:
lm(wage ~ educ, data=mydata)
```


## Regression Output Using stargazer
```{r}
lm.1 <- lm(wage ~ educ, data=mydata) 
stargazer(lm.1,type="text")

bhat <- coef(lm.1)   # create "bhat" as a vector of the estimated coefficients
                     # to make it easier to refer to individual 
                     # estimated coefficients inline
```


The coefficient on years of education (*educ*) is **`r bhat["educ"]`**. 

**Interpretation**: for every
additional year of education that a person completes, their hourly wage 
increases on average by **\$`r round(bhat["educ"],2)`**, or an increase
of **\$`r 2000 * round(bhat["educ"],2)`** annually (assuming a typical year
of work is 2,000 hours).

## Your Turn ##############################################################
Re-do the analysis above, but restrict the sample to people who have at least
12 years of education. What differences do you notice?

## Explore the data
```{r}
# Restrict the sample
mydata <- mydata %>% filter(educ >= 12) # sample of HSG and above only

stargazer(mydata,type="text")
gf_histogram(~wage,data=mydata)
gf_histogram(~educ,data=mydata)
gf_point(wage ~ educ,data=mydata)
cor(mydata$wage,mydata$educ)
```


## Regression
```{r}
lm.2 <- lm(wage ~ educ, data=mydata %>% 
             filter(educ >= 12)) 
stargazer(lm.1,lm.2,type="text",
	title = "Does the effect of education differ after HS?",
	column.labels = c("Full Sample", "12 Years+ Educ Only"),
	dep.var.labels = ("Hourly Wage"),
	covariate.labels = ("Yrs of Education"))

bhat.lm2 <- coef(lm.2)   
```

Restricting the sample to high school graduates and above, we
find that the coefficient on years of education (*educ*) is **`r bhat.lm2["educ"]`**. 

**Interpretation**: for every
additional year of education that a person completes *beyond high school*, their hourly wage 
increases on average by **\$`r round(bhat.lm2["educ"],2)`**, or an increase
of **\$`r 2000 * round(bhat.lm2["educ"],2)`** annually (assuming a typical year
of work is 2,000 hours).
