---
title: "Econ 343 Ch 10 R Lecture"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_collapsed: yes
    toc_float: yes
    toc_levels: 3
  beamer_presentation: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#
# ===>> SETUP - this r-chunk does not usually need to be changed by the student.
#
# The following packages cover all of the commands that will be used in this class,
# as well as the files from the Wooldridge textbook.
# 
# For any assignment, you simply start by making a copy of this file and add the R chunks and
# text as needed below this setup chunk. 
#
# Note that these packages have been carefully selected to create a toolkit that is ideally
# suited for economic and business analysis. You might find it useful to keep this template and
# set of packages for future data analysis projects.
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.
library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()
library(scales)     # to format inline r as currency
library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values
library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)   # limits number of decimals to 4 when printing
knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile
```

# R commands for Time series

- Define Yearly time series beginning in yyyy
  - tsdata <- ts(dsname, start=yyyy)
  
- `dynlm()` is time Series version of `lm()`
  
- lags in regression model:

  - lm <- dynlm(y ~ x1 + L(x1) + L(x1,2) + x2, data=tsdata)
  
- trends in regression model
  - trend(tsdata)


# Example 10-4 Lags and LRP

```{r}

# Define Yearly time series beginning in 1913
tsdata <- ts(fertil3, start=1913)

# Linear regression of model with lags:
lm.ex10.4 <- dynlm(gfr ~ pe + L(pe) + L(pe,2) + ww2 + pill, data=tsdata)
stargazer(lm.ex10.4,type = "text")


# F test. H0: all pe coefficients are=0
linearHypothesis(lm.ex10.4, matchCoefs(lm.ex10.4,"pe"))

```


## Calculate the LRP

```{r}
b<-coef(lm.ex10.4)
LRP <- b["pe"]+b["L(pe)"]+b["L(pe, 2)"]
```

The long run propensity impact of the personal tax exemption is: 
**$LRP_{pe} = `r LRP`$**.


## F test. H0: LRP=0
```{r}
linearHypothesis(lm.ex10.4,"pe + L(pe) + L(pe, 2) = 0")
```



# Example 10-7 Trends

## Define Yearly time series beginning in 1947
```{r}
tsdata <- ts(hseinv, start=1947)
as_tibble(tsdata)
```


## Linear regression with and w/o trend:
```{r}
lm.ex10.7 <- dynlm(log(invpc) ~ log(price)                , data=tsdata)
lm.ex10.7t <- dynlm(log(invpc) ~ log(price) + trend(tsdata), data=tsdata)

stargazer(lm.ex10.7,lm.ex10.7t, type="text")
```


