---
title: "Econ 343 Midterm 1 Review"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_levels: 1
    number_sections: false
---

```{r setup, include=FALSE}
#
# Always include these packages in R projects for this class
#
library(readr)      # for reading in .csv files
library(dplyr)      # for filter and select commands. etc.

library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula

library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests
library(lmtest)     # post-reg tests
library(effects)    # to plot partial effects for one regressor
library(visualize)  # to visualize hypothesis test p-values

library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data

options(digits=4)
options(scipen=999)

knitr::opts_chunk$set(echo = TRUE,cache=FALSE)
```


# The Five Steps of Economic Analysis

1. Start with an **Economic Question**.

2. Provide an **Economic Model** that identifies the variable you
want to explain (the dependent variable), and lists the primary explanatory variable(s) and (optionally) additional control variable(s). 

3. Specify the **Econometric Model** by specifying a linear equation
that represents the relationship between the variables. 

4. Estimate the regression model to obtain the 
**estimated regression coefficients** on the explanatory variables ($\hat{\beta_x}$s) and interpret what they mean.

5. Use post-regression analysis to test hypotheses and make predictions.

## Example


### 1. Economic Question

#### How much does the size and number of bedrooms of a house affect its price?

### 2. Economic Model

$$price = f(size, bedrooms, location, location, location)$$

### 3. Econometric Model

$$price = \beta_0 + \beta_1 sqrft + \beta_2 bdrms + u $$

### 4. Estimated Regression

To examine this question, house price, sq. ft, number of bedrooms, and other 
information was gathered on homes from the real estate pages of the *Boston Globe* newspaper in 1990. The dataset *hprice1* accompanies the 
Wooldridge textbook and is loaded with the wooldridge package.

```{r}
mydata <- hprice1
stargazer(as.data.frame(mydata), type = "text")

```


#### House price (in thousands) is regressed on sq. footage and number of bedrooms. 

```{r}

# ==> Run the regression and display the results
lm1 <- lm(price ~ bdrms + sqrft ,data = mydata)
stargazer(lm1,type = "text",digits = 2)

# ==> Save the estimated coefficients from the regression
bhat1   <- coef(lm1)   # ALL COEFFICIENTS are saved in the VECTOR "bhat1"
bhat_0  <- bhat1["(Intercept)"]  # ONE coeffcient from the vector
bhat_sf <- bhat1["sqrft"]        # ONE coeffcient from the vector
bhat_bd <- bhat1["bdrms"]        # ONE coeffcient from the vector

```

#### The estimated regression equation is:

$$\widehat{price} =  `r round(bhat_0,2)` + `r round(bhat_sf,2)` sqrft +
`r round(bhat_bd,2)` bdrms $$

#### Interpretation: 

- Each additional bedroom increases the expected price of a 
house by **\$`r round(bhat_bd,2)`** (thousands), holding size constant.

- Each additional square foot added to the size of a house a size, 
increases the expected price of a house by 
**\$`r round(bhat_sf,2)`** (thousands), 
holding the number of bedrooms constant. 

- Thus, for every additional 
**100 s.f.** the expected price of the house increases by 
**\$`r format(round(bhat_sf * 100 * 1000,2),big.mark=",")`**.


### 5. Post-regression Analysis

#### An 1,800 s.f., three bedroom house is listed for \$266,000. Is this a good deal?

```{r}
listprice  <- 266000
predict_1  <- (bhat_0 + bhat_sf*1800  + bhat_bd * 3) * 1000
residual_1 <- listprice - predict_1

```

- The expected price of a 1,800 s.f., 3 bedroom house is 
**\$`r format(round(predict_1,2),big.mark=",")`** which
is 
**\$`r format(round(residual_1,2),big.mark=",")`**
**less** than the list price. 

- Thus, this house is **overpriced** relative to similar homes its size.

# Gauss-Markov Assumptions

If the G-M assumptions hold, then the estimates from a linear regression model
are BLUE: **B**est **L**inear **U**nbiased **E**stimates

1. Model is linear *in parameters*.
2. The data are from a random sample.
3. The explanatory variables are not perfectly collinear.
4. The model does not omit any relevant variables. (Zero Conditional Mean)
5. The error terms have the same variance across the range of x's. (Homoskedasticity)

# Estimate a regression model

## Economic Question: How does horsepower affect gas mileage?

Use the "lm" command in R to estimate a linear regression model. Use
the stargazer() command to display the results. The following
example uses the *mtcars* dataset that comes with R.

```{r}
lm1 <- lm(mpg ~ hp + wt, data=mtcars)
stargazer(lm1,type="text")
```



# Save the estimated coefficients and $R^2$

``` {r}
# ==> change "lm1" below to the name used to save the regression 
model  <- lm1   

# ==> the following variable is a VECTOR
bhat   <- coef(lm1)          # all estimated coefficients

# ==> the following variables are single VALUES
bhat_0 <- bhat["(Intercept)"]
bhat_hp <- bhat["hp"]
bhat_wt <- bhat["wt"]
r_sqr      <- summary(lm1)$r.squared

```

# Use the saved regression results

- "bhat" is a VECTOR, a list that contains all estimated coefficients

- To obtain the value of a single estimated coefficient from the vector, 
put the name of the variable associated with the coefficient
in double quotes inside square brackets [" "] after bhat, e.g. **bhat["hp"]**.

# Use "inline" *R* in your write-ups

- Inline R is used in the middle of a sentence to run an R command

- The most common uses of inline R are to display an estimated
coefficient, the result of some other statistical command, or a calculation

- Results and calculated values from R chunks should always be obtained using R inline, never hardcoded (typed directly into the sentence). 

# Write out the econometric and estimated models

- The relationship between horsepower on gas mileage is estimated by regression *mpg* on horsepower (*hp*) on *mpg*: 

$$mpg = \beta_0 + \beta_1 hp + \beta_2 wt + u$$

- The estimated model is:
$$\widehat{mpg} = `r bhat["(Intercept)"] ` + `r bhat["hp"] ` hp + `r bhat["wt"] ` wt$$.


- Interpretation: The estimated coefficient on *hp* is 
**`r bhat["hp"] `**. 
An increase of 100 horsepower will reduce gas mileage on average by
**`r abs(round(100*bhat["hp"],1))`** miles per gallon.

- Interpretation: The estimated coefficient on *wt*(in 1,000 lbs) is 
**`r bhat["wt"] `**. 
An increase in weight by 1,000 lbs. will **reduce** gas mileage on average by
**`r abs(round(bhat["wt"],1))`** miles per gallon.


# Omitted Variable Bias

## What happens if an important variable is left out of the model?

- To illustrate what happens to the estimated coefficient 
when there is an omitted variable, suppose that the true model is $mpg = \beta_0 + \beta_1 hp + \beta_2 wt + u$, as above, but the 
estimated model does not include **wt**.

- Unlike in real life, for this illustration we have the data 
on the missing variable. 

- We run two regressions in addition to the estimated regression: 

1. the estimated model 
2. the true model 
3. the omitted variable regressed on the explanatory variable.

```{r}
lm.est  <- lm(mpg ~ hp, data=mtcars)
lm.true <- lm(mpg ~ hp + wt, data=mtcars)
lm.omit <- lm(wt  ~ hp,data=mtcars)
stargazer(lm.est,lm.true,lm.omit, type="text")
```


## Estimating the bias term

Use the regression results to calculate the elements for the bias equation, 
$E[\tilde{\beta}_{hp}] = \hat{\beta}_{hp} + \hat{\beta}_{wt}*\tilde{\delta}_{hp}$

where:

- $\hat{\beta}_{wt}$ = the coefficient on **"wt"** from **lm.true**. This is the correlation between the omitted variable (**wt**) and the dependent variable **mpg**. It is **negative** as expected. Heavier cars get lower gas mileage.

- $\tilde{\delta}_{hp}$ = the coefficient on **"hp"** from **lm.omit**. This is the correlation between the omitted variable (**wt**) and the included variable **hp**. It is **positive** as expected. More powerful cars are typically heavier than other cars.

Because $\hat{\beta}_{wt}$ is negative and $\tilde{\delta}_{hp}$ is positive, the bias term is **negative**. The negative impact of **hp** on 
miles per gallon is underestimated. 

### Conclusion: 
Because the bias is negative and the coefficient is negative, 
the estimate from the
underspecified model suggests that HP pulls down *mpg* **even more** 
than it actually does. 
In the true model where the weight of the car is controlled for, HP does not reduce gas mileage as severely.

