---
title: "R Stats Tutorial"
subtitle: "ECON 343"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggformula)
library(mosaic)
library(wooldridge)

theme_set(theme_bw())

# knitr::opts_chunk$set(exercise.checker = checkr::checkr_tutor)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.width = 6, fig.height = 2.5)
tutorial_options(exercise.eval = FALSE)
```


<style type="text/css">
  span.boxed {
    border:5px solid gray;
    border-radius:10px;
    padding: 5px;
  }
span.invboxed {
  border:5px solid gray;
  padding: 5px;
  border-radius:10px;
  color: white;
}

table, td, th { border:0px; }

/* cellpadding */
  th, td { padding: 5px; }

</style>

## Your Mission
The purpose of this tutorial is to help you start to get familiar with the way R works, and some basic R commands.

But you may have noticed that *you are not even using R.* At least, you haven't logged in to the RStudio server.

This tutorial environment lets you read some helpful information, then practice writing and running your own R code, *without* tabbing back and forth between the tutorial website and RStudio. Here's hoping it provides a nice, gentle introduction in a controlled environment!

Tutorials can also serve as helpful reference guides later on, if you need reminders or more practice. We will keep using them to learn and practice R throughout the semester, and links to all our tutorials will be on the course Moodle page ("R Tutorials" section).

## Communicating with R
  You will do most of your work in R with *code* or *commands*. Instead of pointing and clicking, you will type one or more lines of code, which R will *execute* (doing the work you have asked it to do). Then, R will return the results of whatever operation you asked it to do - sometimes producing a plot, other times creating a plot. Sometimes executing code has almost no visible effect (no plot or text output is produced), but instead some object is created and stored in R's *environment* for later use.

  
### Two Key Questions

To get R (or any software) to do something for you,
there are two important questions you must be able to answer.  Before
continuing, think about what those questions might be.
  
### The Questions

To get R (or any software) to do a job for you, there are two
important questions you must be able to answer:
  
#### 1. What do you want the computer to do?
  
#### 2. What must the computer know in order to do that?
  
### Providing R with the information it needs
  
  R *functions* provide R with the answer to the first question: what do you want the computer to do?
  
  Most functions in R have short, but descriptive names that describe what they do.  For example, R has some functions to do basic mathematical operations:  the function `sqrt()` computes the square root of a number, and the function `round()` rounds a number (by default, it rounds to the nearest integer).
  
  But just giving R a function is not enough: you also need to answer the second question (what information does R need to do the job?). For example, if you want to use the function `round()`, you also need to provide R with the number you want to round!
  
  We will provide answers to our two questions by filling in
the boxes of a basic template:
  
  <center>
  <h2><strong><span class="boxed">function</span> ( <span class="boxed">&nbsp;information1&nbsp; , &nbsp;information2&nbsp; , ...)</strong> 
  </h2></center>
  &nbsp;

(The `...` indicates that there may be some additional *arguments* (input information we could provide to R) we could add
  eventually. Some functions need only one input, but if a function takes more than one argument, they are separated by commas.)

###Using simple functions
Let's practice what you just learned, trying out the `sqrt()` and `round()` functions. 

Edit the code below to compute the square root of 64:

```{r find-sqrt, exercise=TRUE}
function(information_R_needs)
```

```{r find-sqrt-hint-1}
sqrt(information_R_needs)
```

```{r find-sqrt-hint-2}
sqrt(64)
```

Now try computing the square root of 44, then rounding it to the nearest integer:

```{r round-sqrt, exercise=TRUE}
function1(information_R_needs)
function2(information_R_needs)
```

```{r round-sqrt-hint-1}
sqrt(information_R_needs)
function2(information_R_needs)
```

```{r round-sqrt-hint-2}
sqrt(44)
function2(information_R_needs)
```

```{r round-sqrt-hint-3}
sqrt(44)
round(information_R_needs)
```

```{r round-sqrt-hint-4}
sqrt(44)
round(sqrt(44)) # yes, you can nest functions. sometimes (not always) a good idea.
```

### Storing information in R: variables
In the last section, you computed the square root of 44 and then rounded it, probably like this:

```{r, echo=TRUE}
sqrt(44)
round(6.63325)
```

But to do that, you probably had to first find the root, make a note of the result, and then provide that number to the `round` function.  What a pain!

A better alternative, if you are computing a value that you will want to use later on, is to store it as a named variable in R.  In the previous example, you might want to store the square root of 44 in a variable called my_root; then you can provide my_root to the `round` function without checking the result of the `sqrt` calculation first:

```{r naming-variable, echo=TRUE}
my_root <- sqrt(44)
round(my_root)
```

Notice that to assign a name to the results of some R code, you use the symbol "**<-**". You can think of it as an **assignment arrow** -- it points from a result toward a name and assigns the name to the result.

Try editing the code to change the name of the variable from my_root to something else, then run your new code:

```{r name-a-variable, exercise=TRUE}
my_root <- sqrt(44)
round(my_root)
```

### What if I have a list of numbers to store?

Sometime you might want to create a variable that contains more than one number.  You can use the function `c` to **concatenate** (link together) numbers into one list:

```{r}
my_fave_numbers <- c(4, 44, 16)
my_fave_numbers
```

(First we stored the list of numbers, calling it my_fave_numbers; then we printed the results to the screen by simply typing the variable name my_fave_numbers).

Try making a list of your three favorite numbers, then using the function `sum` to add them all up:

```{r c-and-sum, exercise=TRUE}
```

```{r c-and-sum-hint, results='hide'}
my_numbers <- c(14, 27, 455)
sum(my_numbers)
```

### What about data that are not numeric?
R can work with categorical data as well as numeric data.  For example, we could create a list of words and store it as a variable if we wanted (feel free to try changing the words if you want):

```{r cat-var, exercise=TRUE}
my_words <- c('RStudio', 'is', 'awesome')
my_words
```

### Datasets in R are called 'data.frames'
- You will typically use datasets for analysis that are similar
to Excel spreadsheets:
    - one column for every variable
    - one row for every observation (person/place/thing for which data were collected)

## Using built-in packages in R

- R *packages* contain additional R commands and datasets you use for specific purposes.
- You have to load the package before you can use it by running the command:
  
```{r library-example, eval=FALSE}
library(packagename) 
```


### Example of loading a package

- Our textbook has a package that accompanies it: `wooldridge`
- Once you have loaded this package, you will have access to all of the datasets used in the examples and homework assignments

Edit the code below to load the `wooldridge` package and then click "run code" to run it.

**(Nothing obvious will happen...this command just gives R permission to access the package, without asking it to actually DO anything.)**

```{r require-package, echo = FALSE, exercise=TRUE}
library()
```

```{r require-package-hint-1}
library(wooldridge)
```

### Viewing a dataset
The wooldridge package includes a dataset called `wage1`.

If you just enter the dataset name (`wage1`) as a command, R will print some of the dataset out to the screen:

```{r check-out-wage1, exercise=TRUE}
wage1
```

You'll see the first 10 rows and the first few variables. You can 
scroll to see the rest of the variables and more of the observations.

Notice that the type of variable is noted below the names: **`<int>`** = integer, **`<dbl>`** = double, for example.


### Describing datasets
- For built-in datasets (that you don't load yourself), you can use
the help command ('?datasetname') to see where the data came from and what the variables represent

Try it out for the Wooldridge ***wage1*** dataset:

```{r check-out-data-1, exercise=TRUE}
library(wooldridge)
?wage1
```

Note that the formatting is a little strange-looking because this is a tutorial, but when you are in Rstudio, it will look nice.

### Summarizing datasets

- Another way to look at a dataset is to see the summary statistics about all of the variables. The `stargazer` package is useful for this. 
- Use the `stargazer` command from the stargazer package to summarize the dataset
- You must add the `type = "text"` option to the command while in the tutorial

Run the code below to see the summary of ***wage1***. 

```{r check-out-data, exercise=TRUE}
library(stargazer)
library(wooldridge)
stargazer(wage1,type = "text")
```

### Try it yourself. 

The `wooldridge` package has another dataset called ***vote1***. Use the help command to learn about what is in the dataset and use *stargazer* to summarize the variables.

```{r vote1,echo = FALSE, exercise=TRUE}

```

```{r vote1-hint-1}
library(wooldridge)
?vote1
```

```{r vote1-hint-2}
library(wooldridge)
library(stargazer)
stargazer(vote1,type = "text")
```


## Reading in data 
You will also get a chance to work with other datasets that are not already loaded in R. You will have to *upload* files to rstudio before you can read them in. You will simply click the upload button and browse to the file on your hard drive.

## Creating your analysis file
Whether you are using a wooldridge dataset or data from some other source, 
you will want to make a copy of the original data to create your own 
version that you can modify for analysis. 

You will use the "**assignment arrow**" symbol ("**`<-`**) when you read in a dataset to make a copy of it for your own use. 


```{r}
mywage1 <- wage1
```

Now you are ready to begin using your dataset to answer questions!

## Review

What have you learned so far?  More than you think!

### Functions in R
You've learned that R code is made up of functions, which are generally named descriptively according to the job they do.  Functions have one or more input arguments, which is where you provide R with all the data and information it needs to do the job.  The syntax for calling a function uses the template:

  <center>
  <h2><strong><span class="boxed">function</span> ( <span class="boxed">&nbsp;information1&nbsp; , &nbsp;information2&nbsp; , ...)</strong> 
  </h2></center>
  &nbsp;
  
  
### Variables in R
You've practiced creating variables in R using `c`, and saving information (or the results of a computation) using the assignment arrow <-.  

### Datasets in R
You've learned how to get some basic information about a dataset and 
how to create your own copy of a dataset to work with.

## Congratulations!
You just completed your first tutorial on R, and wrote some of your own R code. I *knew* you could do it...

© 2020 GitHub, Inc.
Terms
Privacy
Security
Status
Help
Contact GitHub
Pricing
API
Training
Blog
About
50 results found.