---
title: "Econ 343 Practice Exam"
output:
  html_document: default
---

**Instructions:** 

- Complete each of the questions below. Use r "mini-chunks" in
the write-ups. Make sure your write-ups are your own words and interpretations.

- The prompt **Write your answer here:                       ** indicates where you should
include your written responses. Do not put written responses elsewhere. Make sure your
answer is in between the two sets of asterisks so that it is bold.

- Display all estimated regression results using stargazer, with type="text". 

- When complete, knit to HTML, open in browser, then print, 
using "Save as PDF" option. Upload the .pdf to Moodle.


```{r setup, include=FALSE}
#
# Always include these packages in R projects for this class
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.

library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()

library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values

library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data


options(digits=4)   # limits number of decimals to 4 when printing

knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile

```


# 1  Economic Question: What is the effect of experience on wages for urban, college-educated workers? 
(a) (10 points) Create a dataset named *mydata* that contains only 
the college educated (educ>=16) people from the ***wage2*** dataset.(All datasets are from the wooldridge package.)
A
graph command is provided for you that 
compares the effect of experience on wages for urban and non-urban workers. After creating
the dataset, run the graph command as is.


```{r}
# Uncomment to run the following commands once you have filled in the missing information ("???").

#mydata <- ???

#gf_point(wage ~ exper|urban, data=mydata) %>%
#  gf_lm()

```


From the graph, what
can you say about the effect of experience on wages for Urban vs. Non-Urban
College Educated workers?

**Write your answer here:                       **


(b) (10 points) With *mydata*, estimate a **semi-log** model for *wage* on *education*,*experience*,*tenure*,
the dummy variable *urban* (already provided for you in the dataset) and
an **interaction term between urban and experience** (you provide). 

Just run the model and display the results. Save the discussion of the results 
for the next parts of this question.

```{r}
# Uncomment to run the following commands once you have filled in the missing information ("???").


#Model.1 <- lm(???, data=mydata)

#stargazer(Model.1, type="text")
```

(c) (10 points) Interpret and discuss the statistical significance of each of the coefficients in the model.

**Write your answer here:                       **




(d) (10 points) Test whether the effect of 1 more year of experience on log wages the same has the effect 
as 1 more year of tenure for this group of college-educated workers. What do you find? Explain.

```{r}
 
```

**Write your answer here:                       **





# 2 Economic Question: Are married applicants more likely to be approved for a mortgage?
Use the variables in the *loanapp* dataset. Note that the dummy variable 
*approve* = 1 if the person 
was approved for a loan, 0 otherwise. The dummy variable *married* = 1 if the person is
married, and 0 otherwise. The variable *dep* contains the total number of dependents 
that the applicant has. For simplicity, assume all dependents
are children.

(a) (10 points) Create a dummy variable called "haskid" to indicate whether the applicant has kids or
not.

```{r}
mydata1 <- loanapp

#mydata1$haskid <- ???
# ???
# ???
```

(b) (10 points) Estimate a linear probability model of *approve* on *married*, *appinc*, your
new variable *haskid*, and **one additional variable of your choice**. Report the results using stargazer.


**NOTE: To run your model, uncomment and modify the lm() command in the r-chunk below that has already
been started for you. DO NOT DELETE the option ***na.action=na.exclude*** or
the model results will not work with the predict() and lht() commands later on.
 

```{r}
# Uncomment to run the following commands once you have filled in the missing commands ("???").

#lm.loana  <- lm(approve ~ married + appinc + haskid + ???,data=mydata1,na.action=na.exclude)

#stargazer(lm.loana,type="text")

```

(c) (10 points) Explain your reasoning behind the additional variable you chose. What did you expect
to learn by including it? Save your discussion about what you did learn for the next part.

**Write your answer here:                       **


(d) (10 points) Interpret the meaning and discuss the statistical significance 
of the coefficients on *married*, *haskid* and *the additional variable you included*. 

**Write your answer here:                       **


(e) (10 points) Does the model result in any unreasonable estimated probabilities? Explain.

```{r}



```

**Write your answer here:                       **

(f) (10 points) What proportion of the observations are correctly predicted? What do you conclude from this? 

``` {r}

```

**Write your answer here:                       **


# 3 Economic Question: Does Spending increase school performance?
Imagine you are one of several interns working for a Michigan legislator. The legislators wants 
to know how much school performance improves with additional per pupil spending. 

Another intern drafted a report that concluded there was no relationship on
spending and academic performance. The report included regression results to support that conclusion, 
using the 2001 MEAP data on 4th grade math scores and per pupil expenditures. 

 
The estimates included in the report were:
$$ \widehat{math4_i} = {74.870} - {0.001}*exppp_i$$  
where *math4* is the percent of 4th graders with
a passing grade on the state-wide math exam for a particular school and
*exppp* is per pupil spending for that school. The coefficient on *exppp* was reported to be statistically
insignificant.

You believe that the model used in the report is underspecified. That is, it has omitted an important variable
from the model. Specifically, you believe that the percent of students from low-income families in a school
(the variable *lunch*)
should be included in the model for the model to be correctly specified. Without *lunch* in the model, the estimated coefficient on *exppp* reported above is likely to be biased.

Your task is to help your fellow intern understand the danger of omitted variable bias. To help you do this,
you complete the following tasks.

(a) First, explain what each of the four parameters in the omitted variable bias equation represent 
for this report: 

$$ E(\widetilde{\beta_1}) = \hat{\beta_1} + \widetilde{\delta_1}\hat{\beta_2}$$

Write the meaning of each of the components below in terms that your fellow intern will understand. 
The meaning should be specific to this
economic question. The meaning should not include a numerical estimate.

**Write your answers below.**

(5 points) **$E(\widetilde{\beta_1})$:          **

(5 points) **$\hat{\beta_1}$:                   **

(5 points) **$\widetilde{\delta_1}$:            **

(5 points) **$\hat{\beta_2}$:                   **

(10 points) Explain in a sentence or two, how the omitted variable bias equation applies to the conclusion the
intern drew from the analysis in the draft report.

**Write your answer here:                       **



Next, run the following three models, and present the results side-by-side in stargazer. Don't discuss the results
until the parts that follow.

(b) (10 points) Replicate the other intern's results using the *meap01* dataset.

(c) (10 points) Run the correct model, that is, add the variable *lunch* to the model.

(d) (10 points) Run the additional model required to explain omitted variable bias.

```{r}

# lm.omitvarbias <- 
# lm.corrected   <- 
# lm.additional  <-

#stargazer(lm.omitvarbias, lm.corrected, lm.additional, type="text",
#         title="put a relevant title here")


```

(e) (10 points) Using the estimated coefficients as needed from the three regressions, 
quantify and explain the
impact of omitted variable bias on the conclusions of the other intern's report.
(you can just type in the actual values of the estimated coefficients rather than 
using r-mini-chunks, to save time)

**Write your answer here:                       **



(f) (10 points) Fill in the following equation to report the estimated results from the correctly specified model. 

$$ \widehat{math4_i} = {???} + {???}*exppp_i + ???*lunch_i$$ 

Write a concluding paragraph that discusses what this analysis can say about the impact of school spending on school performance.

**Write your answer here:                       **

