---
title: "Econ 343 Practice Exam Solution"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_collapsed: yes
    toc_float: yes
    toc_levels: 3
---


```{r setup, include=FALSE}
#
# Always include these packages in R projects for this class
#
library(readr)      # for reading in .csv files
library(tidyverse)  # data wrangling
library(dplyr)      # for filter and select commands. etc.

library(wooldridge) # datasets in the textbook
library(stats)      # basic statistic commands
library(DataExplorer) # nice commands to learn about a new data set
library(stargazer)  # nice regression output
library(ggformula)  # easy graphing interface for ggplot
library(mosaic)     # accompanies ggformula
library(expss)      # apply_labels() cro()

library(AER)        # data and commands specific to econometrics
library(car)        # for linearhypothesis tests lht(), predict() hccm() heterosced.
library(lmtest)     # post-reg tests coeftest()
library(effects)    # to plot effects of one regressor, holding all others constant
library(visualize)  # to visualize hypothesis test p-values

library(dynlm)      # lm commands for time series data
library(plm)        # lm commands for panel data


options(digits=4)   # limits number of decimals to 4 when printing

knitr::opts_chunk$set(echo = TRUE) # TRUE = commands in r chunks to be echoed in outfile

```
Instructions: 

- Complete each of the questions below. You do not need to use r "mini-chunks" in
the write-ups. Make sure your write-ups are your own words and interpretations.

- Display all estimated regression results using stargazer, with type="text". 
Do not use the option results="asis" at the top of the r chunk. 

180 points total. 

# 1  Wages and Experience
**Economic Question: What is the effect of experience on Wages for Urban, College Educated workers? **

(a) (10 points) Create a dataset named *mydata* that contains only 
the college educated (educ>=16) people from the wage2 dataset. A
graph command is provided for you that 
compares the effect of experience on wages for urban and non-urban workers. Run
the graph command as is.

**From the graph, what
can you say about the effect of experience on wages for Urban vs. Non-Urban
College Educated workers?**

```{r}
mydata <- wage2 %>% filter(educ >= 16)

gf_point(wage ~ exper|urban, data=mydata) %>%
  gf_lm()

```


(b) (10 points) With *mydata*, estimate a **semi-log** model for *wage* on *education*,*experience*,*tenure*,
the dummy variable *urban* (already provided for you in the dataset) and
an **interaction term between urban and experience** (you provide). 

Just run the model and display the results. Save the discussion of the results 
for the next parts of this question.

```{r}
# ANSWER


Model.1 <- lm(log(wage) ~ educ + exper + tenure + urban + urban*exper, data=mydata)

stargazer(Model.1, type="text")
```

(c) (10 points) Interpret and discuss the statistical significance of each of the coefficients in the model.

**ANSWER: Wages in urban areas are on average 21.6% higher than in non-urban areas. One additional
year of experience in urban areas reduces wages, on average, by .3%, however
the estimate is not statistically significant.**

(d) (10 points) Test whether the effect of 1 year of experience on log wages the same has the effect 
as 1 more year of tenure for this group of college-educated workers. What do you find? Explain.

```{r}
# ANSWER
lht(Model.1,"exper=tenure")
```

**ANSWER: They do appear to have different effects. We fail to reject that the coefficients are
equal to each other.**

# 2 Mortgage approval and Married
**Economic Question: Are married applicants more likely to be approved for a mortgage?**

Use the variables in the *loanapp* dataset. Note that the dummy variable 
*approve* = 1 if the person 
was approved for a loan, 0 otherwise. The dummy variable *married* = 1 if the person is
married, and 0 otherwise. The variable *dep* contains the total number of dependents 
that the applicant has.

(a) (10 points) Create a dummy variable called "haskid" to indicate whether the applicant has kids or
not.

(b) (10 points) Estimate a linear probability model of *approve* on *married*, *appinc*, your
new variable *haskid*, and **one additional variable of your choice**. Report the results using stargazer.


**NOTE: To run your model, uncomment and modify the lm() command in the r-chunk below that has already
been started for you. I added the option ***na.action=na.exclude*** to it just in case the variable you choose to add to the model is missing for any observations. You need to leave that option in there or
the model results will not work with the lht() command later on.
 

```{r}

mydata <- loanapp

mydata$haskid <- NULL
mydata$haskid[mydata$dep == 0] = 0
mydata$haskid[mydata$dep >  0]= 1

lm.loana  <- lm(approve ~ married + appinc + haskid,data=mydata,na.action=na.exclude)

stargazer(lm.loana,type="text")

```

(c) (10 points) Explain your reasoning behind the additional variable you chose. What did you expect
to learn by including it? Save your discussion about what you did learn for the next part.


(d) (10 points) Interpret the meaning and discuss the statistical significance 
of the coefficients on *married*, *haskid* and *the additional variable you included*. 



(e) (10 points) Does the model result in any unreasonable estimated probabilities? Explain.

```{r}

mydata$yhat <- fitted(lm.loana)
count(mydata$yhat<0)
count(mydata$yhat>1)

```

(f) (10 points) What proportion approvals are correctly predicted by this model? What do you conclude from this?

```{r}

mydata$yhat <- fitted(lm.loana)
favstats(mydata$yhat)

mydata$yhat_dummy <- NULL
mydata$yhat_dummy[mydata$yhat>=.5] = 1 
mydata$yhat_dummy[mydata$yhat <.5] = 0 

mydata$corr_pred <- NULL
mydata$corr_pred[mydata$yhat_dummy == mydata$approve] = 1
mydata$corr_pred[mydata$yhat_dummy != mydata$approve] = 0

favstats(mydata$corr_pred)
```

(g) (10 points) What proportion of married and unmarried applicants are correctly predicted? What do you conclude from this? 
``` {r}
favstats(mydata$corr_pred[mydata$married==1])
favstats(mydata$corr_pred[mydata$married==0])
```


# 3 Spending and School Performance
**Economic Question: Does Spending increase school performance?**

Imagine you are one of several interns working for a Michigan legislator. The legislators wants 
to know how much school performance improves with additional per pupil spending. 

Another intern drafted a report that concluded there was no relationship on
spending and academic performance. The report included regression results to support that conclusion, 
using the 2001 MEAP data on 4th grade math scores and per pupil expenditures. 

 
The estimates included in the report were:
$$ \widehat{math4_i} = {74.870} - {0.001}*exppp_i$$  
where *math4* is the percent of 4th graders with
a passing grade on the state-wide math exam for a particular school and
*exppp* is per pupil spending for that school. The coefficient on *exppp* was reported to be statistically
insignificant.

You believe that the model used in the report is underspecified. That is, it has omitted an important variable
from the model. Specifically, you believe that the proportion of students from low-income families in a school
(the variable *lunch*)
should be included in the model. Without it, the estimated coefficient reported above is biased.

Your task:
(a) (10 points) Replicate the other intern's results.
(b) (10 points) Run the correct model,
(c) (10 points) Run an additional model needed to explain omitted variable bias.
(d) (10 points) Present the results of all three models side-by-side in stargazer.
(e) (10 points) Explain what the equation for omitted variable bias is and what each of the four parameters in
the equation stand for: 
$$ E(\widetilde{\beta_1}) = \hat{\beta_1} + \widetilde{\delta_1}\hat{\beta_2}$$
(f) (10 points) Using the results from the three regressions and the omitted variable bias equation, quantify and explain the
impact of omitted variable bias on the conclusions of the other intern's report.

```{r}

mydata <- meap01

lm1 <- lm(math4 ~ exppp,data=mydata)
lm2 <- lm(lunch ~ exppp, data=mydata)
lm3 <- lm(math4 ~ exppp + lunch, data=mydata)
stargazer(lm1,lm2,lm3,type="text",
          title="Estimating Omitted Variable Bias")


```

(g) (10 points) Write a concluding paragraph (or a set of bullet points) that discusses what this analysis can say about
the impact of school spending on school performance.





